---
title: "Analysis of soil carbon controls across Washington State"
author: "Anthony Stewart"
output:
  html_document:
    df_print: paged
    theme: readable
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

h1, h3, h4 {
  text-align: center;
}
```

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.show = "hold", time_it = TRUE, dpi = 100)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = T, collapse = TRUE)
knitr::opts_knit$set(root.dir = '/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/')
library(rgl)
library(terra)
library(lme4)
library(MASS)
library(mgcv)
library(lmerTest)
library(MuMIn)
library(RLRsim)
library(terra)
library(spatialEco)
library(sf)
library(mapview)
library(car)
library(ggplot2)
library(forcats)
library(sjPlot)
library(sjstats)
library(DHARMa)
library(ggeffects)
library(merTools)
library(glmnet)
library(stats)
library(ggcorrplot)
library(RColorBrewer)
library(cowplot)
library(webshot)
library(kableExtra)
library(pdp)
library(vip)
library(formatR)
library(dplyr)
library(ggpubr)
library(grid)
library(gridExtra)

knitr::knit_hooks$set(webgl = hook_webgl)
rgl::setupKnitr(autoprint = TRUE)
```

```{r include=FALSE}

# Quick R^2 function
r.sq <- function(y,y.fitted){
    res <- y-y.fitted
    1-sum(res^2)/sum((y-mean(y))^2)
}
```


```{r echo=FALSE}
hoh_dat <- data.frame(vect("SOIL CARBON/All_WA/data/points/hoh_pts_2855.gpkg"))
mas_dat <- data.frame(vect("SOIL CARBON/All_WA/data/points/mas_pts_2856.gpkg"))
col_dat <- data.frame(vect("SOIL CARBON/All_WA/data/points/col_pts_2855.gpkg"))

wa_dat <- rbind(hoh_dat, mas_dat, col_dat) |> 
  mutate(
  GEO = as.factor(GEO),
  geomorphons = as.factor(geomorphons),
  site = as.factor(site),
  site = fct_reorder(site, SOC_stock_spline, .fun = "median")) |>
  dplyr::rename_with(~gsub("_median", "", .x, fixed = TRUE))
wa_dat
```

```{r}
unique(hoh_dat$GEO)
unique(mas_dat$GEO)
unique(col_dat$GEO)
```


With many potential predictors we can eliminate some by examining correlation plots

```{r echo=FALSE}
wa_dat_num <- wa_dat |> 
    dplyr::select(SOC_stock_spline,
               #lower_depth,
               DTM,
               HLI,
               # NDYI_median, # Normalized difference yellow index
               # NDVI_median, # Normalized difference vegetation index
               #MNDWI, #Modified Normalized Difference Water Index
               EVI, # Enhanced vegetation index
               WIP) |> as.matrix()

ggcorrplot(cor(wa_dat_num), method = "square", type = "full", lab = T, lab_size = 3)

```

There is weak correlation between `SOC_stock_spline` and the selected covariates. Additionally there is some collinearity between predictors. The spectral predictors such as NDYI (Normalized Difference Yellow Index) and NDVI (Normalized Difference Vegetation Index). We can remove these based on the correlation coefficient > 0.7.

Now we can begin examining the distribution of carbon stock values in the dataset to choose the appropriate transformation. We also scale and center the numeric predictor variables.

```{r echo=FALSE, fig.show='hold', out.width="35%"}

columns_to_exclude <- c("SOC_stock_spline", "lower_depth") 

wa_dat_scale <- wa_dat |> 
    dplyr::select(sample_ID, lower_depth, SOC_stock_spline, 
                  site, DTM, geomorphons, GEO, WIP,
                  EVI, MNDWI, HLI) |> 
    dplyr::mutate(across(
        dplyr::where(is.numeric) & !all_of(columns_to_exclude),
                  ~dplyr::case_when(TRUE ~ scale(.))),
        site = fct_reorder(site, SOC_stock_spline, .fun = "median")) 
(wa_dat_scale)
hist(wa_dat_scale$SOC_stock_spline)
hist(log10(wa_dat_scale$SOC_stock_spline))

```

### Explicit parameter model building 


Now build models using log transformed carbon stock data. We need to specify that `sample_ID` is a random effect because of the multiple samples at one location. `lower_depth` will be a random slope to adjust model based on how it is affected by depth. 

Also lower depth is included as a fixed effect to specify an overall effect of depth in addition to accounting for the random slope within `sample_ID` locations. See:

- [https://stats.stackexchange.com/questions/614812/can-random-slopes-also-be-included-as-fixed-effects](https://stats.stackexchange.com/questions/614812/can-random-slopes-also-be-included-as-fixed-effects)

We also use a different optimizer for the lmer model which uses `nloptwrap` by default and gives convergence warnings for multiple model iterations. We use the `bobyqa` optimizer and have assessed negligiable changes to log liklihood (< 0.1) and parameter estimates (< 0.00001). See:
- [https://cran.r-project.org/web/packages/lme4/vignettes/lmerperf.html](https://cran.r-project.org/web/packages/lme4/vignettes/lmerperf.html)

- [https://stackoverflow.com/questions/33670628/solution-to-the-warning-message-using-glmer](https://stackoverflow.com/questions/33670628/solution-to-the-warning-message-using-glmer) and `help(convergence)`

We could use a Generalized Linear Model or a Generalized Linear Mixed Model here too but they often fail to converge. We will model with the log-transformed soil carbon stocks in the linear mixed model 

```{r}
#Full model with all parameters
mod1 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP+EVI+DTM + (GEO) + geomorphons + site +
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
# No geomorphonssum
mod2 <-lmer(log10(SOC_stock_spline) ~ 
                WIP+EVI+DTM+HLI + (GEO) + site +
                 lower_depth + (lower_depth|sample_ID),  
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#No spectral, geomorphons
mod3 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP+DTM+HLI + (GEO) + site +
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#No geology or geomorphons
mod4 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP+DTM+HLI + site +
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#No WIP, geomorphons
mod5 <- lmer(log10(SOC_stock_spline) ~ 
                  DTM+HLI + (GEO) + site +
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#HLI interaction within site
mod6 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP + DTM +(GEO) + HLI*site +
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#DTM interaction with site
mod7 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP + HLI + DTM*site + (GEO) + 
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#Just geology and WIP
mod8 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP + (GEO) + 
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#Just WIP
mod9 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP + 
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#NULL
mod10 <- lmer(log10(SOC_stock_spline) ~ 
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))


```

Pairwise comparisons between the top, largest model and the rest show if models with fewer parameters are significantly different. Considering AIC will help determine the most parsimonious model with the best fit to the data. 

```{r tidy=FALSE, echo=FALSE}
anova_table_mod1 <- rbind(anova(mod1, mod2), 
                            anova(mod1, mod3), 
                            anova(mod1, mod4), 
                            anova(mod1, mod5),
                            anova(mod1, mod6), 
                            anova(mod1, mod7), 
                            anova(mod1, mod8), 
                            anova(mod1, mod9), 
                            anova(mod1, mod10)) 

anova_table_mod1 |>
    tibble::rownames_to_column(var = "models") |> 
    dplyr::mutate(Significant = case_when(`Pr(>Chisq)` < 0.05 & `Pr(>Chisq)`>  0.01~ "***",
                                `Pr(>Chisq)` < 0.01 & `Pr(>Chisq)` > 0.001~ "***",
                                `Pr(>Chisq)` < 0.001 ~ "***",
                                .default = ""),
           Models = case_when(models == "mod11" ~ "mod1",
                              models == "mod12" ~ "mod1",
                              models == "mod13" ~ "mod1",
                              models == "mod14" ~ "mod1",
                              models == "mod15" ~ "mod1",
                              models == "mod16" ~ "mod1",
                              models == "mod17" ~ "mod1",
                              models == "mod18" ~ "mod1",
                              .default = models)) |> 
    dplyr::select(-models) |> 
    dplyr::select(Models, everything()) |> 
    #dplyr::arrange((AIC)) |> 
    kbl() |>
    #row_spec(, background = "lightblue") |> 
    kable_classic_2("hover", full_width = F) 
```
From the ANOVAs it looks like `model 7` has the lower AIC when compared together. 


To hone in on differences between variations in `model 7` we can remove some of the variables and compare AICs for a more parsimonious fit.

```{r}
#DTM interaction with site
#No HLI
mod7.1 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP  + (GEO) + DTM:site +
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#No GEO, HLI
mod7.2 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP + DTM:site +
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
#No WIP
mod7.3 <- lmer(log10(SOC_stock_spline) ~ 
                 DTM:site +
                 lower_depth + (lower_depth|sample_ID), 
            data = wa_dat_scale, REML = F,
            control=lmerControl(optimizer="bobyqa"))
```



```{r tidy=FALSE, echo=FALSE}
anova_table_mod2 <- rbind(anova(mod7, mod7.1), #Not Significant
anova(mod7, mod7.2), 
anova(mod7, mod7.3), 
make.row.names = T) #Not Significant

anova_table_mod2 |>
    tibble::rownames_to_column(var = "models") |> 
    mutate(Significant = case_when(`Pr(>Chisq)` < 0.05 & `Pr(>Chisq)`>  0.01~ "***",
                                `Pr(>Chisq)` < 0.01 & `Pr(>Chisq)` > 0.001~ "***",
                                `Pr(>Chisq)` < 0.001 ~ "***",
                                .default = ""),
           Models = case_when(models == "mod71" ~ "mod7",
                              models == "mod72" ~ "mod7",
                              models == "mod73" ~ "mod7",
                              .default = models)) |>
    dplyr::select(-models) |> 
    dplyr::select(Models, everything()) |> 
    kbl() |>
    kable_classic_2("hover", full_width = F)

```
`Model 7` is the lowest AIC

 The model $R^2$ for `mod7` is `r round(r.sq(log10(wa_dat_scale$SOC_stock_spline), fitted(mod7)), 3)`

```{r include=FALSE}
round(r.sq(log10(wa_dat_scale$SOC_stock_spline), fitted(mod7)), 3)
```


Confidence intervals calculated with bootstrapping show that `WIP` and  `Precip` are the most significant contributors to SOC stock. the overlapping confidence intervals do not show much separation between `WIP` and  `Precip` but there is more variation in the `Precip` confidence interval.

<div align="center">
```{r cache=TRUE, echo=FALSE, out.width="100%"}
tab_model(mod7, bootstrap = T, seed = 11, dv.labels = c("Model 7"))
```
</div>

Looking at some partial dependence plots we can see some relationships between the predicted values and `WIP` and `Precip`. The pattern in `Precip` shows the differences between the two study areas. with wetter areas having a higher range of `SOC stock` 

```{r fig.show='hold', fig.align='center', echo=FALSE, message=FALSE}
library(ggpubr)

partial_plot <- function(data, model, x_var, color_var, shape_var){
  if(is.factor(data[x_var][[1]]) == TRUE) {
  x_var <- sym(x_var)
  shape_var <- sym(shape_var)
  color_var <- sym(color_var) 
  fitted <- predict(model, newdata = data)
  
  ggplot(data, aes(y = 10**fitted, x = as.factor(!!x_var))) +
    geom_violin(show.legend = FALSE, scale = "width", linewidth = 0.9)+
    geom_jitter(aes(shape = !!shape_var, , color = !!color_var), width = 0.2, size = 2)+
    xlab("Study Area\n") + 
    ylab(expression('Fitted Model SOC Stock (g cm'^-2*')')) +
    theme(legend.position = 'right', 
          legend.key.size = unit(0.6, "cm"),
          legend.spacing.x = unit(1.2, "cm"),
          panel.background = element_blank(),
          panel.grid.major = element_line(colour = "grey80"),
          axis.ticks = element_blank(),
          text = element_text(size = 16))
  } else {
    if(x_var == "WIP"){
      label <- "Wetland Intrinsic Potential \n Scaled"
    } else if(x_var == "DTM") {
      label <- "Elevation Scaled\n"
    }
  x_var <- sym(x_var)
  shape_var <- sym(shape_var)
  color_var <- sym(color_var) 
  fitted <- predict(model, newdata = data)
  
  ggplot(data, aes(y = 10**fitted, x = !!x_var)) +
    geom_point(aes(shape = !!shape_var, color = !!color_var), size = 2)+
    xlab(label) +
    ylab(expression('Fitted Model SOC Stock (g cm'^-2*')')) +
    geom_smooth(aes(y = 10**fitted, x = !!x_var), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                linewidth = 0.9, linetype = 5, alpha = 0.3, se = T) + 
    theme(legend.position = 'top', 
          legend.key.size = unit(0.8, "cm"),
          legend.spacing.x = unit(1.2, "cm"),
          panel.background = element_blank(),
          panel.grid.major = element_line(colour = "grey80"),
          axis.ticks = element_blank(),
          text = element_text(size = 16))
  }
}

mod7site <- partial_plot(data = wa_dat_scale, model = mod7, x_var = "site", shape_var = "GEO", color_var = "HLI")

mod7wip <- partial_plot(data = wa_dat_scale, model = mod7, x_var = "WIP", shape_var = "GEO", color_var = "HLI")

mod7dtm <- partial_plot(data = wa_dat_scale, model = mod7, x_var = "DTM", shape_var = "GEO", color_var = "HLI")

ggarrange(mod7site, mod7wip+rremove("ylab"), mod7dtm+rremove("ylab"), ncol = 3, nrow=1, common.legend = T,
          vjust = 10)
```

#### Splitting up wetland and upland soils 

Use the WIP threshold of 50% or 0.5

```{r echo=FALSE, message=FALSE}
wa_dat_scale_wet <- wa_dat |> 
    filter((WIP >= 0.5 )) |> 
    dplyr::mutate(across(
        dplyr::where(is.numeric) & !all_of(columns_to_exclude),
                  ~dplyr::case_when(TRUE ~ scale(.))))
    
wa_dat_scale_upl <- wa_dat |> 
    filter((WIP < 0.5 )) |> 
    dplyr::mutate(across(
        dplyr::where(is.numeric) & !all_of(columns_to_exclude),
                  ~dplyr::case_when(TRUE ~ scale(.))))
    
```

<div align="center">
```{r echo=FALSE, message=FALSE}
mod7_wet <- update(mod7, data = wa_dat_scale_wet)

mod7wet_pred <- predict(mod7, newdata = wa_dat_scale_wet)
#summary(mod2.1_wet)
mod7_upl <- update(mod7, data = wa_dat_scale_upl)

mod7upl_pred <- predict(mod7, newdata = wa_dat_scale_upl)
#summary(mod2.1_upl)
tab_model(mod7_wet, mod7_upl, dv.labels = c("Wetlands (WIP > 0.50", "Uplands (WIP < 0.50)"))
```
</div>

```{r echo=FALSE, message=FALSE}
library(ppcor)

rbind(pcor.test(wa_dat_scale_wet$WIP, wa_dat_scale_wet$SOC_stock_spline, wa_dat_scale_wet$lower_depth),
pcor.test(wa_dat_scale_upl$WIP, wa_dat_scale_upl$SOC_stock_spline, wa_dat_scale_upl$lower_depth))

rbind(pcor.test(wa_dat_scale_wet$DTM, wa_dat_scale_wet$SOC_stock_spline, wa_dat_scale_wet$lower_depth),
pcor.test(wa_dat_scale_upl$DTM, wa_dat_scale_upl$SOC_stock_spline, wa_dat_scale_upl$lower_depth))

rbind(pcor.test(wa_dat_scale_wet$HLI, wa_dat_scale_wet$SOC_stock_spline, wa_dat_scale_wet$lower_depth),
pcor.test(wa_dat_scale_upl$HLI, wa_dat_scale_upl$SOC_stock_spline, wa_dat_scale_upl$lower_depth))

```


```{r echo=FALSE, message=FALSE}
library(ggpubr)

DTM_wet <- partial_plot(data = wa_dat_scale_wet, 
             model = mod7, 
             x_var = "DTM", 
             color_var = "site", 
             shape_var = "GEO")

wip_wet <- partial_plot(data = wa_dat_scale_wet, 
             model = mod7, 
             x_var = "WIP", 
             color_var = "site", 
             shape_var = "GEO")

DTM_upl <- partial_plot(data = wa_dat_scale_upl, 
             model = mod7, 
             x_var = "DTM", 
             color_var = "site", 
             shape_var = "GEO")

wip_upl <- partial_plot(data = wa_dat_scale_upl, 
             model = mod7, 
             x_var = "WIP", 
             color_var = "site", 
             shape_var = "GEO")

wetupl_graphs <- ggarrange(DTM_wet+rremove("xlab"), wip_wet+rremove("xlab")+rremove("ylab"), 
                           DTM_upl, wip_upl+rremove("ylab"), ncol = 2, nrow=2, common.legend = T)
annotate_figure(
    annotate_figure(wetupl_graphs, 
                    left = grid::textGrob("Upland", 
                                      rot = 90, vjust = 0.5, hjust = 3, 
                                      gp = grid::gpar(cex = 1.3))), 
    left = grid::textGrob("Wetland", 
                    rot = 90, vjust = 2.1, hjust = -1.5, gp = grid::gpar(cex = 1.3)))

ggplot2::ggsave(plot = wetupl_graphs, filename = paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/mod2.1_wet_upl.jpeg"),
                width = 9, height = 7.5, units = "in", dpi = 500)

```


We can visualize some of these interactions with interaction plots against SOC stock. 

```{r fig.show='hold', message=FALSE, echo=FALSE}
library(interactions)
DTMsite <- interactions::interact_plot(mod7, pred = DTM, modx = site, plot.points = T, plot.shape = T,
                            x.label = "DTM", 
                            y.label = "Fitted Model SOC Stock",
                            legend.main = "Site",
                            interval = TRUE, )


DTMsite
```


### Random Forest Machine Learning


Using Random Forest can see if a more flexible, machine learning model can capture any non-linear relationships or interaction relationships between parameters.

`lower_depth` is included as a predictor here in the data setup

```{r message=FALSE, echo=FALSE}
library(randomForest)
set.seed(11)

# Validation Set 
train.index <- as.vector(sample(c(1:nrow(wa_dat_scale)), 0.7*nrow(wa_dat_scale), replace=F))
train <- wa_dat_scale[train.index, -1]
train$GEO <- as.character(train$GEO)
train$SOC_stock_spline <- log10(train$SOC_stock_spline)

test <- wa_dat_scale[-train.index, -1]
test$GEO<- as.character(test$GEO)
test$SOC_stock_spline <- log10(test$SOC_stock_spline)

full <- wa_dat_scale[, -1]

```


We then use the `tuneRF` to choose the appropriate `mtry` number 

```{r echo=FALSE, message=FALSE, results='hide'}
set.seed(11)
tuneRF(x = train[, 4:ncol(train)], y = (train$SOC_stock_spline), mtryStart = 4, ntreeTry = 500, stepFactor = 1.5, improve = 0.0001)
```


```{r fig.show='hold', cache=TRUE, results='hide'}
set.seed(11)
rf_model <- randomForest((SOC_stock_spline) ~ .,
                         sampsize = nrow(train),
                         ntree = 1000, mtry = 4,
                         importance = TRUE, data = train)
rf_model
plot(rf_model)

rf.train <- predict(rf_model, newdata = train)
rf.test <- predict(rf_model, newdata = test)

vip::vip(rf_model)
```


61.63% of the out of bag (OOB) variation is explained. Looks like `lower_depth`, `WIP`, and `site` are the big drivers. 


```{r echo=FALSE}

errdf <- data.frame("error variable" = c("MAE", "MAE^2", "RMSE Test", "RMSE non-log", "Stdev of non-log SOC stock",  "R^2 on test dataset"), 
                    "value" = c(caret::MAE(pred = rf.train, obs = (train$SOC_stock_spline)),
                                caret::MAE(pred = rf.train, obs = (train$SOC_stock_spline))**2,
                                caret::RMSE(pred = rf.test, obs = (test$SOC_stock_spline)),
                               exp(caret::RMSE(pred = rf.test, obs = (test$SOC_stock_spline))),
                                sd(wa_dat_scale$SOC_stock_spline),
                                caret::R2(pred = rf.test, obs = (test$SOC_stock_spline))))
errdf |> kbl() |>
  kable_classic_2("hover", full_width = F)
```


Looking at partial dependency plots we can see how `Precip` and `NDYI` have higher influence on separating variation in model predictions. `WIP` less so but seems to play a role on the higher end. 


```{r tidy=FALSE, echo=FALSE, message=FALSE, cache=TRUE}
library(gridExtra)
library(grid)


wip_pdp<- data.frame(randomForest::partialPlot(rf_model, train, WIP, plot = FALSE)) |> 
    ggplot() +
        geom_line(aes(x = x, y = y)) +
    xlab("WIP") +
    ylim(-1, -0.25) +
    ylab(expression('Model Prediction Effect'))
DTM_pdp<- data.frame(randomForest::partialPlot(rf_model, train, DTM, plot = FALSE)) |> 
    ggplot() +
        geom_line(aes(x = x, y = y))+
    xlab("DTM Elevation") +
    ylim(-1,  -0.25) +
    ylab(expression('Model Prediction Effect'))
lower_pdp <- data.frame(randomForest::partialPlot(rf_model, train, lower_depth, plot = FALSE)) |> 
    ggplot() +
        geom_line(aes(x = x, y = y))+
    xlab("Depth") +
    ylim(-1,  -0.25) +
    ylab(expression('Model Prediction Effect'))



grid.arrange(wip_pdp, lower_pdp, DTM_pdp, ncol = 3, nrow=1)
```

We can now visualize our candidate model 7 in a predicted vs. actual plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(RColorBrewer)
(mod7graph <- ggplot(wa_dat, 
                      aes(y = 10**(fitted(mod7)), x = (SOC_stock_spline))) +
    geom_jitter(color='black', 
                aes(fill = (wa_dat$WIP*100), 
                    shape = as.factor(site)),
               size = 4, stroke = 0.9, alpha = 0.8) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlGnBu"),
                         name = "Wetland \nProbability %", n.breaks = 5, limits = c(0, 100)) +
     scale_shape_manual(name = "Geologic Age", 
                        values = c(21, 22, 23, 24)) +
    geom_smooth(aes(y = 10**(fitted(mod7)), x = (SOC_stock_spline)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                linewidth = 0.9, linetype = 5, alpha = 0.3, se = T) +
    xlab(expression('Sampled SOC Stock (g cm'^-2*')')) + 
    ylab(expression('Predicted SOC Stock (g cm'^-2*')')) + 
    geom_abline(intercept = 0, slope = 1, linewidth = 0.9, linetype = "dashed") +
    annotate("text", 
             label = paste("R^{2} == ", 
                           round(r.sq((wa_dat$SOC_stock_spline),
                                      10**fitted(mod7)), 3)), 
             x = 0.5, y = 3.5, size = 4, parse = T) + 
    annotate("text", label = "Model Fit", 
             x = 1.5, y = 3.5, size = 4) +
    annotate("segment", color = "#fa3e3e",
             x = 1.3, y = 3.3, linewidth = 0.9, linetype = 5, 
             xend = 1.75) + 
    annotate("text", label = "1:1", 
             x = 2.5, y = 3.5, size = 4) +
    annotate("segment", color = "black",
             x = 2.3, y = 3.3, linewidth = 0.9, linetype = "dashed",
             xend = 2.75) +
    xlim(0, 4) +
    ylim(0, 4)  +
    theme(legend.position = 'right', 
          legend.key.size = unit(0.6, "cm"),
          legend.spacing.x = unit(1.2, "cm"),
          panel.background = element_blank(),
          panel.grid.major = element_line(colour = "grey80"),
          axis.ticks = element_blank(),
          text = element_text(size = 16)) +
    guides(guide_legend(byrow = TRUE))
)


```


```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
ggplot2::ggsave(plot = mod7graph, filename = paste0("/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/model7_nonlog_SOCprediction.png"),
                width = 9, height = 7.5, units = "in", dpi = 500)
```


