---
title: 'Data processing for geospatial predictors'
author: "Anthony Stewart"
output: html_document
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

h1, h4, h3 {
  text-align: center;
}
```

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.show = "hold", time_it = TRUE, dpi = 75)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = T, collapse = TRUE)
knitr::opts_knit$set(root.dir = '/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/')
library(terra)
library(aqp)
library(ithir)
library(mpspline2)
library(dplyr)
library(spatialEco)
library(whitebox)
library(stringr)

#setGDALconfig("GDAL_PAM_ENABLED", "FALSE")

terraOptions(
    memfrac = 0.1
)
```

### Raster and Vector data

#### Additional terrain variables

In particular,

-   DEM elevation is a proxy for local climate variation
-   Heat Load Index (HLI) As an estimate of solar radiation intesity
-   landform classification from `geomorphons` is derived using **Whitebox Tools** which implements the algorithm from *Jasiewicz, J., and Stepinski, T. F. (2013). Geomorphons — a pattern recognition approach to classification and mapping of landforms. Geomorphology, 182, 147-156.*
-   

```{r}
terrain_metrics <- function(DEM, path, studyarea){
  stopifnot(file.exists(paste0(path, studyarea, "_HLI", ".tif")) ==FALSE)
  stopifnot(file.exists(paste0(getwd(), "/", path, studyarea, "_geomorphons", ".tif")) == FALSE)
  
  hli <- spatialEco::hli(DEM, force.hemisphere = "northern")
  writeRaster(hli, names = "HLI", 
              paste0(path, studyarea, "_HLI", ".tif"),
              overwrite = TRUE)
  #geomorphons
  whitebox::wbt_geomorphons(
    dem = sources(DEM),
    output = paste0(getwd(), "/", path, studyarea, "_geomorphons", ".tif"),
    search = 100
  ) 
    
}
```

```{r}
path <- "SOIL CARBON/All_WA/data/Rasters/"
col_dem <- rast("SOIL CARBON/All_WA/data/Rasters/Colville_dem_huc2855.tif")
mas_dem <- rast("SOIL CARBON/All_WA/data/Rasters/Mashel_DEM_HUC12_2856_4m.tif")
hoh_dem <- rast("SOIL CARBON/All_WA/data/Rasters/hoh_dem_2855.tif")
```


```{r eval=FALSE}
terrain_metrics(col_dem, path, "Colville")
terrain_metrics(mas_dem, path, "Mashel")
terrain_metrics(hoh_dem, path, "Hoh")
```

#### Other Geospatial Predictors

```{r}
col_studyarea <- vect("SOIL CARBON/All_WA/data/Vectors/ColvilleHUC_2855.shp") 
mas_studyarea <- vect("SOIL CARBON/All_WA/data/Vectors/Mashel_HUC12_2856.shp")
hoh_studyarea <- vect("SOIL CARBON/All_WA/data/Vectors/HOH_POLYGON_711.shp")
```

Geology from the Washington Dept. of Natural Resources 

```{r eval=FALSE}
wa_geology <- vect("WA_Geo/WA_Geology_100K.gpkg") |> tidyterra::select(GEOLOGIC_AGE)

crop(wa_geology, project(hoh_studyarea, wa_geology)) |> 
  project("EPSG:2855") |> 
  writeVector("SOIL CARBON/All_WA/data/Vectors/Hoh_geology.shp")

crop(wa_geology, project(mas_studyarea, wa_geology)) |> 
  project("EPSG:2856") |> 
  writeVector("SOIL CARBON/All_WA/data/Vectors/Mashel_geology.shp")

crop(wa_geology, project(col_studyarea, wa_geology)) |>
  project("EPSG:2855") |> 
  writeVector("SOIL CARBON/All_WA/data/Vectors/Colville_geology.shp")

```


Consolidate the levels in Geologic age to be more broad because our samples did not cover some categories 

- Quaternary
  - Quaternary
  - Present 
- Pleistocene
  - Pleistocene
- Miocene
  - Miocene, 
  - Miocene-Oligocene
  - Miocene, middle to upper
- Eocene
  - Miocene-Eocene
  - Eocene
  - Oligocene-Eocene
- PreTertiary
  - Traissic-Permian
  - pre-Tertiary
  - Permian



```{r eval=FALSE}
hoh_consolidate_geo <- vect("SOIL CARBON/All_WA/data/Vectors/Hoh_geology.shp") |> 
  tidyterra::mutate(GEO = case_when(GEOLOGIC_A == "Miocene-Eocene" ~ "Eocene",
                                    GEOLOGIC_A == "Miocene" ~ "Miocene",
                                    GEOLOGIC_A == "Present" ~ "Quaternary",
                                    .default = GEOLOGIC_A))
mas_consolidate_geo <- vect("SOIL CARBON/All_WA/data/Vectors/Mashel_geology.shp") |> 
  tidyterra::mutate(GEO = case_when(GEOLOGIC_A == "Miocene-Eocene" ~ "Eocene",
                                    GEOLOGIC_A == "Miocene" ~ "Miocene",
                                    GEOLOGIC_A == "Eocene" ~ "Eocene",
                                    GEOLOGIC_A == "Miocene, middle to upper" ~ "Miocene",
                                    GEOLOGIC_A == "Miocene-Oligocene" ~ "Miocene",
                                    GEOLOGIC_A == "Oligocene-Eocene" ~ "Miocene",
                                    GEOLOGIC_A == "Present" ~ "Quaternary",
                                    .default = GEOLOGIC_A))
col_consolidate_geo <- vect("SOIL CARBON/All_WA/data/Vectors/Colville_geology.shp") |> 
  tidyterra::mutate(GEO = case_when(GEOLOGIC_A == "Traissic-Permian" ~ "PreTertiary",
                                    GEOLOGIC_A == "Miocene" ~ "Miocene",
                                    GEOLOGIC_A == "Eocene" ~ "Eocene",
                                    GEOLOGIC_A == "pre-Tertiary" ~ "PreTertiary",
                                    GEOLOGIC_A == "Permian" ~ "PreTertiary",
                                    GEOLOGIC_A == "	Oligocene-Eocene" ~ "Miocene",
                                    GEOLOGIC_A == "Present" ~ "Quaternary",
                                    .default = GEOLOGIC_A))

writeVector(hoh_consolidate_geo, "SOIL CARBON/All_WA/data/Vectors/Hoh_geology_consolidate.shp",overwrite = TRUE)
writeVector(mas_consolidate_geo, "SOIL CARBON/All_WA/data/Vectors/Mas_geology_consolidate.shp",overwrite = TRUE)
writeVector(col_consolidate_geo, "SOIL CARBON/All_WA/data/Vectors/Col_geology_consolidate.shp",overwrite = TRUE)
```


```{r}
hoh_allgeolevels <- vect("SOIL CARBON/All_WA/data/Vectors/Hoh_geology_consolidate.shp") |>
   tidyterra::select(GEO)
mas_allgeolevels <- vect("SOIL CARBON/All_WA/data/Vectors/Mas_geology_consolidate.shp")|>
   tidyterra::select(GEO) 
col_allgeolevels <- vect("SOIL CARBON/All_WA/data/Vectors/Col_geology_consolidate.shp")|>
   tidyterra::select(GEO)

```


Create Rasters from Geology

```{r eval=TRUE}

hoh_geoage_rast <- vect("SOIL CARBON/All_WA/data/Vectors/Hoh_geology_consolidate.shp") |>
  terra::rasterize(y = hoh_dem, field = "GEO", touches = TRUE, filename = "SOIL CARBON/All_WA/data/Rasters/Hoh_geology.tif", overwrite = TRUE)

mas_geoage_rast <- vect("SOIL CARBON/All_WA/data/Vectors/Mas_geology_consolidate.shp")|>
  terra::rasterize(y = mas_dem, field = "GEO", touches = TRUE, filename = "SOIL CARBON/All_WA/data/Rasters/Mashel_geology.tif", overwrite = TRUE)

col_geoage_rast <- vect("SOIL CARBON/All_WA/data/Vectors/Col_geology_consolidate.shp") |>
  terra::rasterize(y = col_dem, field = "GEO", touches = TRUE, filename = "SOIL CARBON/All_WA/data/Rasters/Colville_geology.tif", overwrite = TRUE)

```


```{r}

hoh_geoage_rast <- rast("SOIL CARBON/All_WA/data/Rasters/Hoh_geology.tif")
mas_geoage_rast <- rast("SOIL CARBON/All_WA/data/Rasters/Mashel_geology.tif")
col_geoage_rast <- rast("SOIL CARBON/All_WA/data/Rasters/Colville_geology.tif")

```

Lithology from the Global Lithological Map

- See a thesis from Jansen 2010

```{r eval=FALSE}

GLiM <- vect("SOIL CARBON/All_WA/data/Vectors/LiMW_GIS 2015.gdb") |> select(Litho) 

col_lithol <- crop(GLiM, project(col_studyarea, "ESRI:54012")) |> project("EPSG:2855") |> writeVector("SOIL CARBON/All_WA/data/Vectors/Colville_lithol.shp")
mas_lithol <- crop(GLiM, project(mas_studyarea, "ESRI:54012")) |> project("EPSG:2856") |> writeVector("SOIL CARBON/All_WA/data/Vectors/Mashel_lithol.shp")
hoh_lithol <- crop(GLiM, project(hoh_studyarea, "ESRI:54012")) |> project("EPSG:2855") |> writeVector("SOIL CARBON/All_WA/data/Vectors/Hoh_lithol.shp")
```


Floodplains

Nardi et al., 2019 GFPLAIN250m, a global high-resolution dataset of Earth’s floodplains

```{r eval=FALSE}
flood <- rast("SOIL CARBON/Floodplain/GFPLAIN TIFF/NA/NA.TIF")

col_flood <- crop(flood, project(col_studyarea, flood), mask = TRUE) |> project("EPSG:2855") |> writeRaster("SOIL CARBON/All_WA/data/Vectors/Colville_flood.tif", overwrite = TRUE)
mas_flood <- crop(flood, project(mas_studyarea, flood), mask = TRUE) |> project("EPSG:2856") |> writeRaster("SOIL CARBON/All_WA/data/Vectors/Mashel_flood.tif", overwrite = TRUE)
hoh_flood <- crop(flood, project(hoh_studyarea, flood), mask = TRUE) |> project("EPSG:2855") |> writeRaster("SOIL CARBON/All_WA/data/Vectors/Hoh_flood.tif", overwrite = TRUE)
```

Spectral metrics from Sentinel-2. The median reflectance over summer months (July-Sept) from 2018-2023 
    -   NDVI
    -   MNDWI
    -   NDYI
    -   EVI
    
```{r eval=FALSE}
spectral_mosaic <- rast("SOIL CARBON/All_WA/data/Rasters/Sentinel2_mosaic.tif")
  
col_spectral <- crop(spectral_mosaic, project(col_studyarea, spectral_mosaic), mask = TRUE) |> project("EPSG:2855") |> writeRaster("SOIL CARBON/All_WA/data/Rasters/Colville_spectral.tif", overwrite = TRUE)
mas_spectral <- crop(spectral_mosaic, project(mas_studyarea, spectral_mosaic), mask = TRUE) |> project("EPSG:2856") |> writeRaster("SOIL CARBON/All_WA/data/Rasters/Mashel_spectral.tif", overwrite = TRUE)
hoh_spectral <- crop(spectral_mosaic, project(hoh_studyarea, spectral_mosaic), mask = TRUE) |> project("EPSG:2855") |> writeRaster("SOIL CARBON/All_WA/data/Rasters/Hoh_spectral.tif", overwrite = TRUE)
```

Canopy Height

```{r fig.show='hold', out.width="30%"}
hoh_chm <- rast("SOIL CARBON/All_WA/data/Rasters/Hoh_CanopyHeight_2855.tif") 
mas_chm <- rast("SOIL CARBON/All_WA/data/Rasters/Mashel_CanopyHeight_2856.tif")
col_chm <- rast("SOIL CARBON/All_WA/data/Rasters/Colville_CanopyHeight_2856.tif")

plot(hoh_chm) 
plot(mas_chm)
plot(col_chm) 
```

Climate data from WorldClim

- Use CHM as reference

```{r}
#| eval: false
prism_ppt <- rast("PRISM_Climate/PRISM_ppt_30yr_normal_800mM4_annual_asc/PRISM_ppt_30yr_normal_800mM4_annual_asc.asc")
prism_temp <- rast("PRISM_Climate/PRISM_tmean_30yr_normal_800mM4_annual_asc/PRISM_tmean_30yr_normal_800mM4_annual_asc.asc")
prism_vpd <- rast("PRISM_Climate/PRISM_vpdmax_30yr_normal_800mM5_annual_asc/PRISM_vpdmax_30yr_normal_800mM5_annual_asc.asc")

prism_clim <- c(prism_ppt, prism_temp, prism_vpd)
names(prism_clim) <- c("MAP", "MAT", "VPD")

#wa_clim_prj <- project(prism_clim, "EPSG:4326")
mas_studyareabuff <- buffer(terra::aggregate(mas_studyarea), 1000)

hoh_clim <- crop(prism_clim, project(hoh_studyarea, "EPSG:4326")) |> 
  project("EPSG:2855") |>
  crop(y = project(hoh_studyarea, "EPSG:2855"), mask = TRUE) |> 
  resample(y = hoh_chm) |> 
  mask(mask = hoh_chm, filename = "SOIL CARBON/All_WA/data/Rasters/hoh_climate_data.tif", overwrite = TRUE)

mas_clim <- crop(prism_clim, project(mas_studyareabuff, "EPSG:4326"), touches = F, extend = TRUE) |> 
  project("EPSG:2856") |>
  crop(y = project(mas_studyarea, "EPSG:2856"), mask = TRUE) |> 
  resample(y = mas_chm) |> 
  mask(mask = mas_chm, filename = "SOIL CARBON/All_WA/data/Rasters/mashel_climate_data.tif", overwrite = TRUE)

col_clim <- crop(prism_clim, project(col_studyarea, "EPSG:4326")) |> 
  project("EPSG:2855") |>
  crop(y = project(col_studyarea, "EPSG:2855"), mask = TRUE) |> 
  resample(y = col_chm) |> 
  mask(mask = col_chm, filename = "SOIL CARBON/All_WA/data/Rasters/colville_climate_data.tif", overwrite = TRUE)
```

```{r}
hoh_clim <- rast("SOIL CARBON/All_WA/data/Rasters/hoh_climate_data.tif")
mas_clim <- rast("SOIL CARBON/All_WA/data/Rasters/mashel_climate_data.tif")
col_clim <- rast("SOIL CARBON/All_WA/data/Rasters/colville_climate_data.tif")
```


#### LANDFIRE existing vegetation type 

From Yu et al., 2021:
- “Eleven vegetation types defined from the National Land Cover Database and provided by NEON were binned into four broad types based on preliminary analyses: 
  - deciduous/evergreen/mixed forest (hereafter forest),
  - grassland-herbaceous/pasture-hay/shrub-scrub/dwarf-scrub (grassland/shrubland), 
  - cultivated crops (cropland), and 
  - emergent herbaceous wetlands/sedge herbaceous/woody wetlands (wetland).” 
  
  ([Yu et al., 2021, p. 4](zotero://select/library/items/Q2EDH8L3)) ([pdf](zotero://open-pdf/library/items/2KX4MEH4?page=4&annotation=3KPFC9NZ))

```{r eval=FALSE}

lf_classes <- read.csv("SOIL CARBON/All_WA/data/dataframes/WA_LandFire_Classes.csv") |> select(-X) |> select(values, names) 
lf_mat <- matrix(c(lf_classes[[1]], lf_classes[[2]]), ncol = 2)
lf_wa <- rast("SOIL CARBON/All_WA/data/WA_GEE_Data/WA_LandFire_2014.tif") |> 
  terra::as.factor()


lf_wa_rcl <- mask(lf_wa, lf_wa >= 3001, maskvalues = FALSE, updatevalue = NA) 
#plot(lf_wa_rcl)

hoh_studyarea_4326 <- project(hoh_studyarea, "EPSG:4326")
mas_studyarea_4326 <- project(mas_studyarea, "EPSG:4326")
col_studyarea_4326 <- project(col_studyarea, "EPSG:4326")

lf_hoh <- crop(lf_wa_rcl, hoh_studyarea_4326, mask = TRUE) |>
  project("EPSG:2855") |>
  resample(y = hoh_chm, method = "near") |>
  mask(mask = hoh_chm, 
       filename = "SOIL CARBON/All_WA/data/Rasters/NonPredictors/hoh_LANDFIRE.tif", overwrite = TRUE)
lf_mas <- crop(lf_wa_rcl, mas_studyarea_4326, mask = TRUE)|>
  project("EPSG:2856") |>
  resample(y = mas_chm, method = "near") |> 
  mask(mask = mas_chm, 
       filename = "SOIL CARBON/All_WA/data/Rasters/NonPredictors/mashel_LANDFIRE.tif", overwrite = TRUE)
lf_col <- crop(lf_wa_rcl, col_studyarea_4326, mask = TRUE) |>
  project("EPSG:2855") |>
  resample(y = col_chm, method = "near") |> 
  mask(mask = col_chm, 
       filename = "SOIL CARBON/All_WA/data/Rasters/NonPredictors/colville_LANDFIRE.tif", overwrite = TRUE)

LF_reclass_func <- function(rastcat, class_df, name){
  reclass_df <- levels(rastcat)[[1]] |> as_tibble() |> 
    left_join(y = class_df, by = join_by(ID == values)) |> select(-EVT) |> 
    mutate(names = case_when(
      str_detect(names, pattern = regex("spruce|hemlock|fir|pine|conifer|Tsuga|larch|Evergreen",
                                        ignore_case = T)) ~ "Conifer",
      str_detect(names, pattern = regex("shrub|sagebrush|Bush", 
                                        ignore_case = T)) ~ "Shrubland",
      str_detect(names, pattern = regex("grass|Savanna|Herbaceous|crop|Pasture|Wheat", 
                                        ignore_case = T)) ~ "Herbaceous",
      str_detect(names, pattern = regex("Oak|Mahogany|Deciduous|Aspen|Orchard", 
                                        ignore_case = T)) ~ "Deciduous",
      str_detect(names, pattern = regex("Mixed|swamp|riparian", 
                                        ignore_case = T)) ~ "Mixed",
      str_detect(names, pattern = regex("Sparse|barren|Volcanic|snow", 
                                        ignore_case = T)) ~ "SparseBarren",
      str_detect(names, pattern = regex("developed|mines", 
                                        ignore_case = T)) ~ "Urban",
      .default = names))
  
  lf_recat <- categories(rastcat, layer=1, value = reclass_df, active = 1)
  coltab(lf_recat) <- NULL
  writeRaster(lf_recat, paste0("SOIL CARBON/All_WA/data/Rasters/", name, "_Landfire_reclass", ".tif"), overwrite = TRUE)
  return(lf_recat)
}

lf_hoh_classes <- LF_reclass_func(lf_hoh, class_df = lf_classes, name = "Hoh")
lf_mas_classes <- LF_reclass_func(lf_mas, class_df = lf_classes, name = "Mashel")
lf_col_classes <- LF_reclass_func(lf_col, class_df = lf_classes, name = "Colville")

plot(lf_hoh_classes)
plot(lf_mas_classes)
plot(lf_col_classes)
```

```{r}
lf_hoh <- rast("SOIL CARBON/All_WA/data/Rasters/Hoh_Landfire_reclass.tif")
lf_mas <- rast("SOIL CARBON/All_WA/data/Rasters/Mashel_Landfire_reclass.tif")
lf_col <-rast("SOIL CARBON/All_WA/data/Rasters/Colville_Landfire_reclass.tif")
```



WIP or The **Wetland Intrinsic Potential** (WIP) is derived from implementing the methodology described in *Halabisky, Meghan, Dan Miller, Anthony J. Stewart, Amy Yahnke, Daniel Lorigan, Tate Brasel, and Ludmila Monika Moskal. “The Wetland Intrinsic Potential Tool: Mapping Wetland Intrinsic Potential through Machine Learning of Multi-Scale Remote Sensing Proxies of Wetland Indicators.” Hydrology and Earth System Sciences 27, no. 20 (October 20, 2023): 3687–99. <https://doi.org/10.5194/hess-27-3687-2023>.*

The  WIP of the Hoh River Watershed is taken from Halabisky et al., 2023. Colville and Mashel WIP models were derived using the same approach and shown in the R-Quarto Document `All_WA_WIP_revision.qmd`

#### Non-predictor geospatial data

Hansen forest disturbance maps. 

- Was there forest disturbance in the last 20 years? 

```{r}

```

National wetland condition assessment (NWCA) mapping dataset from Uhran et al., 2021

```{r eval=FALSE}
CONUS_NWCA_Mean <- rast("NWCA Data/CONUS_Stock_Nov_20/CONUS_Full_Stock_Mean.tif")
CONUS_NWCA_Max <- rast("NWCA Data/CONUS_Stock_Nov_20/CONUS_Full_Stock_Max.tif")
CONUS_NWCA_Min <- rast("NWCA Data/CONUS_Stock_Nov_20/CONUS_Full_Stock_Max.tif")

NWCA_extract <- function(CONUS_NWCA_raster, studyarea, ref_raster, path, type){
  name <- str_extract(deparse(substitute(studyarea)), "hoh|mas|col")
  type <- tools::toTitleCase(type)
  
  studyarea_prj <- project(studyarea, CONUS_NWCA_raster)
  
  NWCA <- crop(CONUS_NWCA_raster, studyarea_prj, mask = TRUE) |>
  project(y = ref_raster, threads = TRUE,
          filename = paste0(path, "NonPredictors/", name, "_NWCA_Full_", type, ".tif"),
          overwrite = TRUE)
  return(NWCA)
}

hoh_NWCA <- NWCA_extract(CONUS_NWCA_Mean, hoh_studyarea, hoh_chm, path, "Mean")
mas_NWCA <- NWCA_extract(CONUS_NWCA_Mean, mas_studyarea, mas_chm, path, "Mean")
col_NWCA <- NWCA_extract(CONUS_NWCA_Mean, col_studyarea, col_chm, path, "Mean")

hoh_NWCA <- NWCA_extract(CONUS_NWCA_Max, hoh_studyarea, hoh_chm, path, "Max")
mas_NWCA <- NWCA_extract(CONUS_NWCA_Max, mas_studyarea, mas_chm, path, "Max")
col_NWCA <- NWCA_extract(CONUS_NWCA_Max, col_studyarea, col_chm, path, "Max")

hoh_NWCA <- NWCA_extract(CONUS_NWCA_Min, hoh_studyarea, hoh_chm, path, "Min")
mas_NWCA <- NWCA_extract(CONUS_NWCA_Min, mas_studyarea, mas_chm, path, "Min")
col_NWCA <- NWCA_extract(CONUS_NWCA_Min, col_studyarea, col_chm, path, "Min")
```



#### Raster Stack Processing

- Resample to resolution of WIP output maps 

```{r}
col_wip <- rast("SOIL CARBON/All_WA/data/Rasters/Colville_WIP_Final_2855.tif")[[2]]
mas_wip <- rast("SOIL CARBON/All_WA/data/Rasters/Mashel_WIP_Final_2856.tif")[[2]]
hoh_wip <- rast("SOIL CARBON/All_WA/data/Rasters/Hoh_WIP_Final_2855.tif")
```

Topo data from WIP 
```{r eval=FALSE, include=FALSE}
topo_resample_write <- function(topo_path, WIP, Watershed, rastpath){
  filelist <- list.files(topo_path, 
                         pattern = "*.tif", full.names = TRUE,
                         ignore.case = TRUE,
                         include.dirs = FALSE) |> 
  str_subset(pattern = "grad|dev") |>
  str_subset(pattern = "_50.tif|_300.tif|_1000.tif") |> 
  str_subset(pattern = "^(?!.*xml)") |> 
  str_subset(pattern = "^(?!.*ovr)")
  
  f <- lapply(filelist, terra::rast)

  rs <- list()
  for (i in 1:length(f)) {
    if(res(f[[i]])[[1]] != res(WIP)[[1]]){
      rs[[i]] <- resample(f[[i]], WIP, method='near',
                         threads=TRUE) 
    } else {
      rs[[i]] <- f[[i]]
    }
  }

  s <- rast(rs) 
  s <- terra::project(s, (WIP), method = "near", filename = paste0(getwd(),
                               "/",
                               rastpath,
                               Watershed,
                               "_TopoStack",
                               ".tif"), overwrite = TRUE)
}
```

```{r}
list.files("Mashel/Mashel_WIP_2024/DataExport", 
                        pattern = "*.tif", full.names = TRUE,
                        ignore.case = TRUE,
                        include.dirs = FALSE) |> 
 str_subset(pattern = "grad|dev") |>
 str_subset(pattern = "_50.tif|_300.tif|_1000.tif") |> 
 str_subset(pattern = "^(?!.*xml)") |> 
 str_subset(pattern = "^(?!.*ovr)")
```


```{r eval=FALSE, include=FALSE}
rastpath <- "SOIL CARBON/All_WA/data/Rasters"

topo_resample_write(topo_path = "WIP/Hoh_data/TopoIndices/", WIP = hoh_wip, Watershed = "Hoh", rastpath = rastpath)

topo_resample_write(topo_path = "Mashel/Mashel_WIP_2024/DataExport/", WIP = mas_wip, Watershed = "Mashel", rastpath = rastpath)

topo_resample_write(topo_path = "Colville/Colville_WIP_2024/DataExport/", WIP = col_wip, Watershed = "Colville", rastpath = rastpath)

```

Predictor stack resampling
```{r}
stack_resampling <- function(Watershed, WIP, path) {
            
  filelist <- list.files("SOIL CARBON/All_WA/data/Rasters", 
                         pattern = Watershed, full.names = TRUE,
                         ignore.case = TRUE,
                         include.dirs = FALSE) |> 
              str_subset(pattern = "^(?!.*flood)") |>
              str_subset(pattern = "^(?!.*aux.xml)") |>
              str_subset(pattern = "^(?!.*Mask.tif)") |>
              str_subset(pattern = "^(?!.*hillshade)") |>
              str_subset(pattern = "^(?!.*slpasp)") |>
              str_subset(pattern = "^(?!.*fullmodel_v08)")
  f <- lapply(filelist, terra::rast)
  
  rs <- list()
  for (i in 1:length(f)) {
    if(res(f[[i]])[[1]] != res(WIP)[[1]]){
      rs[[i]] <- resample(f[[i]], WIP, method='near', 
                         threads=TRUE)
    } else if(crs(f[[i]], describe = TRUE)$code != crs(WIP, describe = TRUE)$code){
      rs[[i]] <- project(f[[i]], WIP, method = "near")
    } else {
      rs[[i]] <- f[[i]]
    }
  }
  
  s <- rast(rs)
  writeRaster(s, paste0(getwd(), 
                               "/", 
                               path, 
                               Watershed, 
                               "_PredictorStack", 
                               ".tif"), overwrite = TRUE)
  return(s)
}

```


```{r}
list.files("SOIL CARBON/All_WA/data/Rasters", 
                         pattern = "Mashel", full.names = TRUE,
                         ignore.case = TRUE,
                         include.dirs = FALSE) |> 
              str_subset(pattern = "^(?!.*flood)") |>
              str_subset(pattern = "^(?!.*aux.xml)") |>
              str_subset(pattern = "^(?!.*Mask.tif)") |>
              str_subset(pattern = "^(?!.*hillshade)") |>
              str_subset(pattern = "^(?!.*slpasp)") |>
              str_subset(pattern = "^(?!.*fullmodel_v08)")
```



```{r eval=FALSE}
predpath <- "SOIL CARBON/All_WA/data/Rasters/PredictorStacks/"
hoh_stack <- stack_resampling("Hoh", hoh_wip, predpath)
mas_stack <- stack_resampling("Mashel", mas_wip, predpath)
col_stack <- stack_resampling("Colville", col_wip, predpath)

```

Review stacks

```{r}
hoh_stack <- rast("SOIL CARBON/All_WA/data/Rasters/PredictorStacks/Hoh_PredictorStack.tif") |>
  tidyterra::rename(DTM = hoh_dtm_resamp_7_23,
                    geomorphons = Hoh_geomorphons,
                    WIP = WET,
                    CHM = Hoh_CanopyHeight) 
writeRaster(hoh_stack, "SOIL CARBON/All_WA/data/Rasters/PredictorStacks/Hoh_PredictorStack_Class.tif", overwrite = TRUE)
  
mas_stack <- rast("SOIL CARBON/All_WA/data/Rasters/PredictorStacks/Mashel_PredictorStack.tif") |> tidyterra::select(-UPL) |>
  tidyterra::rename(DTM = Band_1,
                    geomorphons = Mashel_geomorphons,
                    WIP = WET,
                    CHM = lewis_2009_dsm_46)

writeRaster(mas_stack, "SOIL CARBON/All_WA/data/Rasters/PredictorStacks/Mas_PredictorStack_Class.tif", overwrite = TRUE)

col_stack <- rast("SOIL CARBON/All_WA/data/Rasters/PredictorStacks/Colville_PredictorStack.tif") |> tidyterra::select(-UPL) |>
  tidyterra::rename(DTM = ColvilleDEM,
                    geomorphons = Colville_geomorphons,
                    WIP = WET,
                    CHM = colville_2008_dsm_0)

writeRaster(col_stack, "SOIL CARBON/All_WA/data/Rasters/PredictorStacks/Col_PredictorStack_Class.tif", overwrite = TRUE)
```


```{r}
hoh_stack <- rast("SOIL CARBON/All_WA/data/Rasters/PredictorStacks/Hoh_PredictorStack_Class.tif")
mas_stack <- rast("SOIL CARBON/All_WA/data/Rasters/PredictorStacks/Mas_PredictorStack_Class.tif")
col_stack <- rast("SOIL CARBON/All_WA/data/Rasters/PredictorStacks/Col_PredictorStack_Class.tif")

names(hoh_stack) == names(mas_stack) & names(hoh_stack)  == names(col_stack)

hoh_stack$GEO <- subst(hoh_stack$GEO, from = c(0,1,2,3), c("Eocene", "Miocene", "Pleistocene", "Quaternary")) 
mas_stack$GEO <- subst(mas_stack$GEO, from = c(0,1,2,3), c("Eocene", "Miocene", "Pleistocene", "Quaternary")) 
col_stack$GEO <- subst(col_stack$GEO, from = c(0,1,2,3), c("Eocene", "Pleistocene", "PreTertiary", "Quaternary")) 
```


### Point Data

Soil pedon locations contain multiple observations per pedon, one for each horizon.

```{r}
wa_dat <- read.csv("SOIL CARBON/All_WA/data/dataframes/wa_spl_datperc_0_15_30_60_100_150_200.csv") |>
    dplyr::select(sample_ID, site, lower_depth, SOC_stock_spline, lat, lon)
str(wa_dat)

wa_dat_pts <- vect(wa_dat, geom = c("lon", "lat"), crs = "EPSG:4326")
plot(wa_dat_pts)
```

### Extraction

The locations of the pedon observations are used to extract all geospatial metrics.

```{r}
wa_dat_pts |> filter(site == "HOH") |> project("EPSG:2855") |> 
  terra::extract(x = hoh_stack, bind = TRUE, method = "simple") |>
  writeVector("SOIL CARBON/All_WA/data/points/hoh_pts_2855.gpkg", overwrite = TRUE)
wa_dat_pts |> filter(site == "MAS") |> project("EPSG:2856") |> 
  terra::extract(x = mas_stack, bind = TRUE) |>
  writeVector("SOIL CARBON/All_WA/data/points/mas_pts_2856.gpkg", overwrite = TRUE)
wa_dat_pts |> filter(site == "COL") |> project("EPSG:2855") |> 
  terra::extract(x = col_stack, bind = TRUE) |>
  writeVector("SOIL CARBON/All_WA/data/points/col_pts_2855.gpkg", overwrite = TRUE)

hoh_pts <- vect("SOIL CARBON/All_WA/data/points/hoh_pts_2855.gpkg")
mas_pts <- vect("SOIL CARBON/All_WA/data/points/mas_pts_2856.gpkg")
col_pts <- vect("SOIL CARBON/All_WA/data/points/col_pts_2855.gpkg")

```

```{r}
data.frame(mas_pts)
```

### Extraction Org Min

```{r}
wa_dat_org_min_pts <- read.csv( "SOIL CARBON/All_WA/data/dataframes/All_WA_horizons_org_min.csv") |> vect(geom = c("lon", "lat"), crs = "EPSG:4326")

wa_dat_org_min_pts |> filter(site == "HOH") |> project("EPSG:2855") |> 
  terra::extract(x = hoh_stack, bind = TRUE) |>
  writeVector("SOIL CARBON/All_WA/data/points/hoh_pts_orgmin_2855.gpkg", overwrite = TRUE)
wa_dat_org_min_pts |> filter(site == "MAS") |> project("EPSG:2856") |> 
  terra::extract(x = mas_stack, bind = TRUE) |>
  writeVector("SOIL CARBON/All_WA/data/points/mas_pts_orgmin__2856.gpkg", overwrite = TRUE)
wa_dat_org_min_pts |> filter(site == "COL") |> project("EPSG:2855") |> 
  terra::extract(x = col_stack, bind = TRUE) |>
  writeVector("SOIL CARBON/All_WA/data/points/col_pts_orgmin__2855.gpkg", overwrite = TRUE)
```

