---
title: 'Data processing for geospatial predictors'
author: "Anthony Stewart"
output: html_document
---
```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

h1, h4, h3 {
  text-align: center;
}
```


```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.show = "hold", time_it = TRUE, dpi = 75)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = T, collapse = TRUE)
knitr::opts_knit$set(root.dir = '/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/')
library(terra)
library(aqp)
library(ithir)
library(mpspline2)
library(dplyr)
```

### Point Data

Soil pedon locations contain multiple observations per pedon, one for each horizon.

```{r}
library(terra)
library(tidyterra)

wa_dat <- read.csv("SOIL CARBON/ANALYSIS/All_WA_horizons_spec_geoage.csv") |>
    dplyr::select(sample_ID, site, depth_cm, BD_g_cm3, carbon_perc, carbon_stock_g_cm2, x, y)
str(wa_dat)

wa_dat_pts <- vect(wa_dat, geom = c("x", "y"), crs = "EPSG:26910")
plot(wa_dat_pts)
```

### Raster and Vector data 

We import and project geospatial data derived from digital elevation models (DEMs) 

In particular, `geomorphons` is derived using **Whitebox Tools** which implements the algorithm from 

*Jasiewicz, J., and Stepinski, T. F. (2013). Geomorphons — a pattern recognition approach to classification and mapping of landforms. Geomorphology, 182, 147-156.*

- The 10 types of geomorphons are given categorical letter codes

```{r eval=FALSE}

catconv <- data.frame(id = 1:10, geomorphons = letters[1:10])

#Colville
col_hli <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/COL/col_HLI.tif") |> project("EPSG:26910")
col_geomorph <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/COL/col_geomorphons_11_23.tif")
col_geomorph_cat <- categories(col_geomorph, layer = 1, value = catconv) |> project("EPSG:26910")


#Mashel
mas_hli <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/MAS/mas_HLI.tif") |> project("EPSG:26910")
mas_geomorph <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/MAS/mas_geomorphons_11_23.tif")|>
    project("EPSG:26910")
mas_geomorph_cat <- categories(mas_geomorph, layer = 1, value = catconv) |> project("EPSG:26910")


#Hoh
hoh_hli <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/HOH/hoh_HLI.tif") |> project("EPSG:26910")
hoh_geomorph <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/HOH/hoh_geomorphons_11_23.tif")
hoh_geomorph_cat <- categories(hoh_geomorph, layer = 1, value = catconv) |> project("EPSG:26910")
```


We merge layers to make extracting geospatial predictors easier

The `merge_wip` refers to the **Wetland Intrinsic Potential** (WIP) tool and is derived from implementing the algorithm in

*Halabisky, Meghan, Dan Miller, Anthony J. Stewart, Amy Yahnke, Daniel Lorigan, Tate Brasel, and Ludmila Monika Moskal. “The Wetland Intrinsic Potential Tool: Mapping Wetland Intrinsic Potential through Machine Learning of Multi-Scale Remote Sensing Proxies of Wetland Indicators.” Hydrology and Earth System Sciences 27, no. 20 (October 20, 2023): 3687–99. https://doi.org/10.5194/hess-27-3687-2023.*


```{r}
#Merge Layers
geomorph_stack <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/ALL_WA_geomorph_merge.tif")#terra::merge(col_geomorph_cat, mas_geomorph_cat, hoh_geomorph_cat, filename = "SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/ALL_WA_geomorph_merge.tif")
hli_stack <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/ALL_WA_HLI_merge.tif")#terra::merge(col_hli, mas_hli, hoh_hli, filename = "SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/ALL_WA_HLI_merge.tif")

merge_wip <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/ALL_WA_WIP_MERGE.tif")
```

Here is a map of the Hoh River Watershed WIP, where `0.0 = 0% Wetland or Upland` and `1.0 = 100% Wetland`

- The gradient captures hydrologic conditions across the landscape and derived mostly from topography metrics at different scales 

```{r echo=FALSE, message=FALSE}
library(rasterVis)

mas_wip <- rast("Mashel/Mashel_WIP_2024/DataExport/Mashel_WIP_Final_2856.tif") #|> project("EPSG:2856")
hoh_wip <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/HOH/Hoh_WIP_Mask0_10_2022.tif") |> project("EPSG:2855")
col_wip <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/COL/colville_NWI_WIP_clip_INV.tif") |> project("EPSG:2855")

colr <- colorRampPalette(RColorBrewer::brewer.pal(9, "YlGnBu"))
levelplot(hoh_wip, margin = F, par.settings=list(
            axis.line=list(col='transparent') # suppress axes and legend outline
          ), scales=list(draw=FALSE), maxpixels = 9e5, 
          col.regions=colr)
```


More geospatial predictors from across WA state. This includes precipitation and temperature from The PRISM Climate Group

*PRISM Climate Group, Oregon State University, https://prism.oregonstate.edu* 



Geology data on geologic age and lithology are downloaded from the WA Dept. of Natural Resources

```{r eval=FALSE}
#All WA layers
wa_state <- vect("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/WA_State_Boundary/WA_State_Boundary.shp") |> project("EPSG:4326")

wa_precip <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/ALL_WA_Precip.tif") |> terra::project("EPSG:26910")
wa_temp <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/ALL_WA_Temp.tif") |> terra::project("EPSG:26910")

wa_geology <- vect("WA_Geo/ger_portal_surface_geology_100k/surface_geology_100k.gdb", layer = "geologic_unit_poly_100k") |> 
    terra::project("EPSG:26910") |> tidyterra::select(LITHOLOGY, GEOLOGIC_AGE)
```

Spectral data are considered to approximate vegetation productivity and health. They are derived from Landsat median reflectance data collected over June-September in the years 2014-2023 and downloaded from Google Earth Engine.

```{r eval=FALSE}
spec_list <- list.files("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/WA_GEE_StudyAreas_Boxes/", pattern = ".tif", full.names = TRUE)

hoh_poly <- vect("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/HOH/HOH_POLYGON_7_11_2022/HOH_POLYGON_711.gpkg") |> project("EPSG:2855")
mas_poly <- vect("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/MAS/MAS_poly/mashel_poly.shp") |> project("EPSG:2856")
col_poly <- vect("Colville/Colville_WIP_2024/DataExport/ColvilleHUC.shp") |> terra::aggregate() |> project("EPSG:2855")
```


```{r eval=FALSE}
rast_list <- list()
for(i in 1:length(spec_list)){
    rast_list[[i]] <- rast(spec_list[[i]]) 
}

hoh_mas_spec2855 <- merge(rast_list[[2]], rast_list[[1]]) 
hoh_box <- ext(-124.647030861616, -123.709196305401, 47.5021021439002, 47.9527148717352)
hoh_spec <- crop(hoh_mas_spec2855, hoh_box) |> project("EPSG:2855") |> mask(mask = hoh_poly, filename = "SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/HOH/Hoh_Spec_Sum2014_2023.tif", overwrite = TRUE)

mas_box <- ext(-122.390044402154, -121.895362438437, 46.7048642408074, 46.9405693599827)
mas_spec <- crop(hoh_mas_spec2855, mas_box) |> project("EPSG:2856") |> mask(mask = mas_poly, filename = "SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/MAS/Mas_Spec_Sum2014_2023.tif", overwrite = TRUE)

col_spec <- rast_list[[3]] |> project("EPSG:2855") |> crop(y = col_poly, mask = TRUE, filename = "SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/COL/Col_Spec_Sum2014_2023.tif", overwrite = TRUE)


```


```{r eval=FALSE}
wa_spec_merge <- terra::merge(wa_spec1, wa_spec2, wa_spec3, filename = "SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/WA_median_GEE_spec_median_merge.tif", overwrite = T) |>
    project("EPSG:26910")
```


```{r}
wa_spec_merge <- rast("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/WA_median_GEE_spec_median_merge.tif")
```

### Extraction

The locations of the pedon observations are used to extract all geospatial metrics. 

```{r}
wa <- readxl::read_xlsx("SOIL CARBON/ANALYSIS/All_PNW_horizons.xlsx")

mas <- wa[wa$site == "MAS",]
hoh_col <- wa[wa$site == "HOH" | wa$site == "COL", ]
mas_pts <- vect(mas, geom = c("x", "y"), crs = "EPSG:26910") |> project("EPSG:2856")
hoh_col_pts <- vect(hoh_col, geom = c("x", "y"), crs = "EPSG:26910") |> project("EPSG:2855")
hoh_pts <- hoh_col_pts[hoh_col_pts$site == "HOH",]
col_pts <- hoh_col_pts[hoh_col_pts$site == "COL",]


```





```{r}

wa_geo_ext <- read.csv("SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/WA_geo_extract.csv")

wa_dat_pts_ext <- wa_dat_pts |> terra::extract(x = geomorph_stack, bind = TRUE, method = "simple") |>
    terra::extract(x = hli_stack, bind = TRUE, method = "simple") |>
    terra::extract(x = wa_precip, bind = TRUE, method = "simple") |> 
    terra::extract(x = wa_temp, bind = TRUE, method = "simple") |>
    terra::extract(x = merge_wip, bind = TRUE, method = "simple") |>
    terra::extract(x = wa_spec_merge, bind = TRUE, method = "simple") |> 
    tidyterra::rename(HLI = lyr.1,
                      Precip = PRISM_ppt_30yr_normal_800mM4_annual_asc,
                      Temp = PRISM_tmean_30yr_normal_800mM4_annual_asc,
                      WIP = WET) |>
    tidyterra::select(-uncertainty) |>
    tidyterra::bind_spat_cols(wa_geo_ext[,c("LITHOLOGY", "GEOLOGIC_AGE")])
(values(wa_dat_pts_ext))
names(wa_dat_pts_ext)

write.csv(wa_dat_pts_ext, file = "SOIL CARBON/ANALYSIS/R/WA_SOC_controls/All_WA_horizons_spec_geoagelith.csv")
writeVector(wa_dat_pts_ext, file = "SOIL CARBON/SPATIAL LAYERS/SPATIAL_LAYERS_7_11_22/All_WA/All_WA_horizons_spec_geoagelith.gpkg", overwrite = T)
```

