---
title: "Analysis of soil carbon controls across the Pacific Northwest"
author: "Anthony Stewart"
output:
  html_document:
    df_print: paged
    theme: readable
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

h1, h3, h4 {
  text-align: center;
}
```

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.show = "hold", time_it = TRUE, dpi = 100)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = T, collapse = TRUE)
knitr::opts_knit$set(root.dir = '/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/')
library(rgl)
library(terra)
library(lme4)
library(MASS)
library(mgcv)
library(lmerTest)
library(MuMIn)
library(RLRsim)
library(terra)
library(spatialEco)
library(sf)
library(mapview)
library(car)
library(ggplot2)
library(sjPlot)
library(sjstats)
library(DHARMa)
library(ggeffects)
library(merTools)
library(glmnet)
library(stats)
library(ggcorrplot)
library(RColorBrewer)
library(cowplot)
library(webshot)
library(kableExtra)
library(pdp)
library(vip)
library(formatR)
library(dplyr)

knitr::knit_hooks$set(webgl = hook_webgl)
rgl::setupKnitr(autoprint = TRUE)
```

```{r include=FALSE}

# Quick R^2 function
r.sq <- function(y,y.fitted){
    res <- y-y.fitted
    1-sum(res^2)/sum((y-mean(y))^2)
}
```


```{r echo=FALSE}
wa_dat <- read.csv("SOIL CARBON/ANALYSIS/All_WA_horizons_spec_geoage.csv")
str(wa_dat)

wa_spl_dat_0_30_60_100 <- read.csv("SOIL CARBON/ANALYSIS/R/WA_SOC_controls/wa_spl_dat_0_30_60_100.csv")
colnames(wa_spl_dat_0_30_60_100) <- gsub("_median","",colnames(wa_spl_dat_0_30_60_100))
str(wa_spl_dat_0_30_60_100)

wa_spl_dat_30_100 <- read.csv("SOIL CARBON/ANALYSIS/R/WA_SOC_controls/wa_spl_dat_30_100.csv") 
colnames(wa_spl_dat_30_100) <- gsub("_median","",colnames(wa_spl_dat_30_100))
#str(wa_spl_dat_30_100)
```

With many potential predictors we can eliminate some by examining correlation plots

```{r echo=FALSE}
wa_spl_dat_num <- wa_spl_dat_0_30_60_100 |> 
    dplyr::select(SOC_stock_spline,
               lower_depth,
               Precip, 
               Temp,
               NDYI,
               NDVI,
               MNDWI,
               EVI,
               SCI,
               DSI,
               LSWI,
               ANDWI,
               SAVI,
               WIP) |> as.matrix()

ggcorrplot(cor(wa_spl_dat_num), method = "square", type = "full", lab = T, lab_size = 3)

```

There is weak correlation between `SOC_stock_spline` and the selected covariates. Additionally there is some collinearity between predictors. The spectral predictors such as EVI (Enhanced Vegetation Index) and SCI (Soil Condition Index) are well correlated with each other and also with NDVI (Normalized Difference Vegetation Index). We can remove these based on the correlation coefficient > 0.7.

Now we can begin examining the distribution of carbon stock values in the dataset to choose the appropriate transformation. We also scale and center the numeric predictor variables.

```{r echo=FALSE, fig.show='hold', out.width="35%"}

columns_to_exclude_spl <- c("SOC_stock_spline", "lower_depth") 

wa_spl_dat_scale <- wa_spl_dat_0_30_60_100 |> 
    dplyr::mutate(across(
        dplyr::where(is.numeric) & !all_of(columns_to_exclude_spl),
                  ~dplyr::case_when(TRUE ~ scale(.))))
str(wa_spl_dat_scale)
hist(wa_spl_dat_scale$SOC_stock_spline)
hist(log10(wa_spl_dat_scale$SOC_stock_spline))

```

### Explicit parameter model building 


Now build models using log transformed carbon stock data. We need to specify that `sample_ID` is a random effect because of the multiple samples at one location. `lower_depth` will be a random slope to adjust model based on how it is affected by depth. We could use a Generalized Linear Model or a Generalized Linear Mixed Model here too but they often fail to converge. We will model with the log-transformed soil carbon stocks in the linear mixed model 

```{r}
#Full model with hypothesized parameters
mod1 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP*Precip*Temp * ANDWI *  NDYI + (GEOLOGIC_AGE) + 
                 lower_depth + (lower_depth||sample_ID), 
            data = wa_spl_dat_scale, REML = F)
# No interactions
mod2 <-lmer(log10(SOC_stock_spline) ~ 
                WIP+Precip+Temp + ANDWI + NDYI + (GEOLOGIC_AGE)+ 
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale, REML = F)
#No spectral
mod3 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP*Precip*Temp + (GEOLOGIC_AGE)+ 
                 lower_depth + (lower_depth||sample_ID), 
            data = wa_spl_dat_scale, REML = F)
#No geology
mod4 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP*Precip*Temp + ANDWI + NDYI + 
                 lower_depth + (lower_depth||sample_ID), 
            data = wa_spl_dat_scale, REML = F)
#No WIP w/ interaction
mod5 <- lmer(log10(SOC_stock_spline) ~ 
                 Precip*Temp + ANDWI + NDYI + (GEOLOGIC_AGE)+
                 lower_depth + (lower_depth||sample_ID), 
            data = wa_spl_dat_scale, REML = F)
#No WIP w/o interaction
mod6 <- lmer(log10(SOC_stock_spline) ~ 
                 Precip+Temp + ANDWI + NDYI +  (GEOLOGIC_AGE)+
                 lower_depth + (lower_depth||sample_ID), 
            data = wa_spl_dat_scale, REML = F)
#No climate
mod7 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP+  + ANDWI + NDYI +  (GEOLOGIC_AGE)+
                 lower_depth + (lower_depth||sample_ID), 
            data = wa_spl_dat_scale, REML = F)
#No climate No spectral
mod8 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP + (GEOLOGIC_AGE)+
                 lower_depth + (lower_depth||sample_ID), 
            data = wa_spl_dat_scale, REML = F)
#Just WIP
mod9 <- lmer(log10(SOC_stock_spline) ~ 
                 WIP + 
                 lower_depth + (lower_depth||sample_ID), 
            data = wa_spl_dat_scale, REML = F)
#NULL
mod10 <- lmer(log10(SOC_stock_spline) ~ 
                 lower_depth + (lower_depth||sample_ID), 
            data = wa_spl_dat_scale, REML = F)


```

**Note** models 5, 8, and 10, failed to converge

Pairwise comparisons between the top, largest model and the rest show if models with fewer parameters are significantly different. Considering AIC will help determine the most parsimonious model with the best fit to the data. 

```{r tidy=FALSE, echo=FALSE}
anova_table_mod1 <- rbind(anova(mod1, mod2), 
                            anova(mod1, mod3), 
                            anova(mod1, mod4), 
                            anova(mod1, mod5),
                            anova(mod1, mod6), 
                            anova(mod1, mod7), 
                            anova(mod1, mod8), 
                            anova(mod1, mod9), 
                            anova(mod1, mod10)) 

anova_table_mod1 |>
    tibble::rownames_to_column(var = "models") |> 
    dplyr::mutate(Significant = case_when(`Pr(>Chisq)` < 0.05 & `Pr(>Chisq)`>  0.01~ "***",
                                `Pr(>Chisq)` < 0.01 & `Pr(>Chisq)` > 0.001~ "***",
                                `Pr(>Chisq)` < 0.001 ~ "***",
                                .default = ""),
           Models = case_when(models == "mod11" ~ "mod1",
                              models == "mod12" ~ "mod1",
                              models == "mod13" ~ "mod1",
                              models == "mod14" ~ "mod1",
                              models == "mod15" ~ "mod1",
                              models == "mod16" ~ "mod1",
                              models == "mod17" ~ "mod1",
                              models == "mod18" ~ "mod1",
                              .default = models)) |> 
    dplyr::select(-models) |> 
    dplyr::select(Models, everything()) |> 
    kbl() |>
    kable_classic_2("hover", full_width = F) 
```
From the ANOVAs it looks like `model 2` & `model 3` performed just as well as `model 1`meaning we should proceed with these. `model 2` removed interactions and `model 3` removed all the spectral metrics but kept interaction. However, `model 2` has 2 less parameters than `model 3` which gives it a lower AIC when compared together. 

```{r echo=FALSE}
anova(mod3, mod2) |> 
    kbl() |>
    kable_classic_2("hover", full_width = F) 
```

To hone in on differences between variations in `model 2` we can remove some of the variables and compare AICs for a more parsimonious fit.

```{r}
#No ANDWI
mod2.1 <- lmer(log10(SOC_stock_spline) ~ 
                WIP+Precip+Temp  + NDYI + (GEOLOGIC_AGE)+ 
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale, REML = F)
#No ANDWI, No NDYI
mod2.2 <- lmer(log10(SOC_stock_spline) ~ 
                WIP+Precip+Temp + (GEOLOGIC_AGE)+ 
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale, REML = F)
#No NDYI
mod2.3 <- lmer(log10(SOC_stock_spline) ~ 
                WIP+Precip+Temp + ANDWI + (GEOLOGIC_AGE)+ 
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale, REML = F)
#No Geology
mod2.4 <- lmer(log10(SOC_stock_spline) ~ 
                WIP+Precip+Temp + ANDWI + NDYI + 
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale, REML = F)
#No NDYI or Geology
mod2.5 <- lmer(log10(SOC_stock_spline) ~ 
                WIP+Precip+Temp + ANDWI + 
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale, REML = F)
#No ANDWI or Geology
mod2.6 <- lmer(log10(SOC_stock_spline) ~ 
                WIP+Precip+Temp + NDYI +  
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale, REML = F)
#No spectral or Geology
mod2.7 <- lmer(log10(SOC_stock_spline) ~ 
                WIP+Precip+Temp +
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale, REML = F)
```

**Note** models 2, 3, 4, 5 failed to converge

```{r tidy=FALSE, echo=FALSE}
anova_table_mod2 <- rbind(anova(mod2, mod2.1), #Not Significant
anova(mod2, mod2.2), #Not Significant
anova(mod2, mod2.3), #Not Significant
anova(mod2, mod2.4), #Not Significant
anova(mod2, mod2.5), #Not Significant
anova(mod2, mod2.6), 
anova(mod2, mod2.7), make.row.names = T) #Not Significant

anova_table_mod2 |>
    tibble::rownames_to_column(var = "models") |> 
    mutate(Significant = case_when(`Pr(>Chisq)` < 0.05 & `Pr(>Chisq)`>  0.01~ "***",
                                `Pr(>Chisq)` < 0.01 & `Pr(>Chisq)` > 0.001~ "***",
                                `Pr(>Chisq)` < 0.001 ~ "***",
                                .default = ""),
           Models = case_when(models == "mod21" ~ "mod2",
                              models == "mod22" ~ "mod2",
                              models == "mod23" ~ "mod2",
                              models == "mod24" ~ "mod2",
                              models == "mod25" ~ "mod2",
                              models == "mod26" ~ "mod2",
                              .default = models)) |>
    dplyr::select(-models) |> 
    dplyr::select(Models, everything()) |> 
    kbl() |>
    kable_classic_2("hover", full_width = F)

```
`Model 2.1` and `2.2` are the lowest AIC and not significantly different from the fit in `Model 2`. Because `Model 2.2` failed to converge, we will move forward with `Model 2.1`. 

Now we can look at the table of all models and compare AICs. We find that Model 2.1 has the lowest AIC followed by ``Model 2.2`` which failed to converge

```{r 'model list and table', message=FALSE, echo=FALSE}
mod_list <- list(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod2.1, mod2.2,
                 mod2.3, mod2.4, mod2.5, mod2.6, mod2.7)
mod_names <- c("mod1", "mod2", "mod3", "mod4", "mod5", "mod6", "mod7", "mod8", "mod9",
               "mod2.1", "mod2.2", "mod2.3", "mod2.4", "mod2.5", "mod2.6", "mod2.7")


mod_AIC = matrix(nrow = length(mod_list), ncol =3)


for (i in 1:length(mod_list)) {
    mod_AIC[i,1] <- round(AIC(mod_list[[i]]), 2)
    mod_AIC[i,2] <- paste0(formula(mod_list[[i]]))[3]
    mod_AIC[i,3] <- mod_names[i]
}
    
delta_AIC <- as.data.frame(cbind(mod_AIC))
delta_AIC$delta <- as.numeric(delta_AIC$V1) -  min(as.numeric(delta_AIC$V1))
names(delta_AIC) <- c("AIC", "formula", "name", "delta")

```

```{r message=FALSE, echo=FALSE, out.width="75%"}
library(kableExtra)
delta_AIC |>
  kbl() |>
  kable_classic_2("hover", full_width = F) |>
    row_spec(c(10), background = "lightblue")
```

 The model $R^2$ for `mod2.1` is `r round(r.sq(log10(wa_spl_dat_scale$SOC_stock_spline), fitted(mod2.1)), 3)`

```{r include=FALSE}
cor(wa_spl_dat_scale$ANDWI, wa_spl_dat_scale$NDYI)
round(r.sq(log10(wa_spl_dat_scale$SOC_stock_spline), fitted(mod2.1)), 3)
round(r.sq(log10(wa_spl_dat_scale$SOC_stock_spline), fitted(mod2.2)), 3)
```


Confidence intervals calculated with bootstrapping show that `WIP` and  `Precip` are the most significant contributors to SOC stock. the overlapping confidence intervals do not show much separation between `WIP` and  `Precip` but there is more variation in the `Precip` confidence interval.

<div align="center">
```{r cache=TRUE, echo=FALSE, out.width="100%"}
tab_model(mod2.1, bootstrap = T, seed = 10, dv.labels = c("Model 2.1"))
```
</div>

Looking at some partial dependence plots we can see some relationships between the predicted values and `WIP` and `Precip`. The pattern in `Precip` shows the differences between the two study areas. with wetter areas having a higher range of `SOC stock` 

```{r fig.show='hold', fig.align='center', echo=FALSE, message=FALSE}
library(ggpubr)
mod2.1precip <- ggplot(wa_spl_dat_scale, aes(y = fitted(mod2.1), x = Precip)) +
    geom_jitter(aes(shape = GEOLOGIC_AGE, color = NDYI))+
    ylab("Fitted Model 2.1")
mod2.1wip <- ggplot(wa_spl_dat_scale, aes(y = fitted(mod2.1), x = WIP)) +
    geom_jitter(aes(shape = GEOLOGIC_AGE, color = NDYI))+
    ylab("Fitted Model 2.1")

ggarrange(mod2.1precip, mod2.1wip, ncol = 2, nrow=1, common.legend = T)
```

```{r echo=FALSE, message=FALSE}
wa_spl_dat_scale_wet <- wa_spl_dat_0_30_60_100 |> 
    mutate(wip_wet = wa_spl_dat_0_30_60_100$WIP) |> 
    filter(wip_wet >= 0.5) |> 
    select(-wip_wet) |> 
    dplyr::mutate(across(
        dplyr::where(is.numeric) & !all_of(columns_to_exclude_spl),
                  ~dplyr::case_when(TRUE ~ scale(.))))
    
wa_spl_dat_scale_upl <- wa_spl_dat_0_30_60_100 |> 
    mutate(wip_wet = wa_spl_dat_0_30_60_100$WIP) |> 
    filter(wip_wet < 0.5) |> 
    select(-wip_wet) |> 
    dplyr::mutate(across(
        dplyr::where(is.numeric) & !all_of(columns_to_exclude_spl),
                  ~dplyr::case_when(TRUE ~ scale(.))))
```

<div align="center">
```{r echo=FALSE, message=FALSE}
mod2.1_wet <- lmer(log10(SOC_stock_spline) ~ 
                WIP + Precip + NDYI + (GEOLOGIC_AGE)+ 
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale_wet, REML = F)
#summary(mod2.1_wet)
mod2.1_upl <- lmer(log10(SOC_stock_spline) ~ 
                WIP+Precip + NDYI + (GEOLOGIC_AGE)+ 
                 lower_depth + (lower_depth||sample_ID),  
            data = wa_spl_dat_scale_upl, REML = F)
#summary(mod2.1_upl)
tab_model(mod2.1_wet, mod2.1_upl, dv.labels = c("Wetlands (WIP > 0.50", "Uplands (WIP < 0.50)"))
```
</div>

```{r echo=FALSE, message=FALSE}
library(ppcor)

pcor.test(wa_spl_dat_scale_wet$WIP, wa_spl_dat_scale_wet$SOC_stock_spline, wa_spl_dat_scale_wet$lower_depth)

pcor.test(wa_spl_dat_scale_upl$WIP, wa_spl_dat_scale_upl$SOC_stock_spline, wa_spl_dat_scale_upl$lower_depth)

pcor.test(wa_spl_dat_scale_wet$Precip, wa_spl_dat_scale_wet$SOC_stock_spline, wa_spl_dat_scale_wet$lower_depth)

pcor.test(wa_spl_dat_scale_upl$Precip, wa_spl_dat_scale_upl$SOC_stock_spline, wa_spl_dat_scale_upl$lower_depth)
```


```{r echo=FALSE, message=FALSE}
library(ggpubr)
precip_wet <- ggplot(wa_spl_dat_scale_wet, 
                     aes(y = (SOC_stock_spline), x = Precip)) +
    geom_jitter(aes(shape = GEOLOGIC_AGE, color = NDYI)) +
    geom_smooth(aes(y = SOC_stock_spline, x = (Precip)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) + 
    ylab(expression('Fitted SOC Stock (g cm'^-2*')'))
wip_wet <- ggplot(wa_spl_dat_scale_wet, 
                     aes(y = SOC_stock_spline, x = WIP)) +
    geom_jitter(aes(shape = GEOLOGIC_AGE, color = NDYI)) +
    geom_smooth(aes(y = SOC_stock_spline, x = (WIP)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) + 
    ylab("")
precip_upl <- ggplot(wa_spl_dat_scale_upl, 
                     aes(y = SOC_stock_spline, x = Precip)) +
    geom_jitter(aes(shape = GEOLOGIC_AGE, color = NDYI)) +
    geom_smooth(aes(y = SOC_stock_spline, x = (Precip)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) + 
    ylab(expression('Fitted SOC Stock (g cm'^-2*')'))
wip_upl <- ggplot(wa_spl_dat_scale_upl, 
                     aes(y = SOC_stock_spline, x = WIP)) +
    geom_jitter(aes(shape = GEOLOGIC_AGE, color = NDYI))+
    geom_smooth(aes(y = SOC_stock_spline, x = (WIP)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) + 
    ylab("")

wetupl_graphs <- ggarrange(precip_wet, wip_wet, precip_upl, wip_upl, ncol = 2, nrow=2, common.legend = T)

annotate_figure(
    annotate_figure(wetupl_graphs, 
                    left = textGrob("Upland", 
                                      rot = 90, vjust = 0.5, hjust = 3, 
                                      gp = gpar(cex = 1.3))), 
    left = textGrob("Wetland", 
                    rot = 90, vjust = 2.1, hjust = -1.5, gp = gpar(cex = 1.3)))

ggplot2::ggsave(plot = wetupl_graphs, filename = paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/mod2.1_wet_upl.jpeg"),
                width = 9, height = 7.5, units = "in", dpi = 500)

```



---

### Dredging for model selection

With the explicit hypothesized approach, we can only evaluate so many models. However, there may be interactions between variables that might be more powerful controls and could give insight into relationships between controls.

The `dredge` function can be used to look through multiple combinations of models from a globally defined model. The number of parameters are limited so we run two models: one with the `NDYI` parameter and another with the `MNDWI` parameter included with `Temp`, `Precip`, and `WIP` fully interacting. `GEOLOGIC_AGE` is also included as an additional, non-interaction parameter and random effects are constant.

```{r, cache=TRUE, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
gmod1 <- lmer(log10(SOC_stock_spline) ~  WIP*Temp*Precip*NDYI+GEOLOGIC_AGE+ lower_depth + (lower_depth||sample_ID), 
              data = wa_spl_dat_scale, REML = F, na.action = 'na.fail')
dredge1 <- dredge(gmod1, beta = "sd" )

```

```{r, cache=TRUE, out.height="50%", warning=FALSE, echo=FALSE, out.width="75%"}
head(dredge1) |>
  kbl() |>
  kable_classic_2("hover", full_width = F) |>
    scroll_box(width = "100%")
dmod1 <- get.models(dredge1, 1)[[1]]
```


```{r cache=TRUE, echo=FALSE, out.width="100%"}
tab_model(dmod1, bootstrap = T, seed = 10, dv.labels = c("Dredge Model"))
```

```{r 'dredge model list and table', message=FALSE, echo=FALSE}
mod_list <- list( get.models(dredge1, 1)[[1]],  get.models(dredge1, 2)[[1]],  get.models(dredge1, 3)[[1]])
mod_names <- c("dmod1", "dmod2", "dmod3" )


mod_AIC = matrix(nrow = length(mod_list), ncol =3)


for (i in 1:length(mod_list)) {
    mod_AIC[i,1] <- round(AIC(mod_list[[i]]), 2)
    mod_AIC[i,2] <- paste0(formula(mod_list[[i]]))[3]
    mod_AIC[i,3] <- mod_names[i]
}
    
delta_AIC <- as.data.frame(cbind(mod_AIC))
delta_AIC$delta <- as.numeric(delta_AIC$V1) -  min(as.numeric(delta_AIC$V1))
names(delta_AIC) <- c("AIC", "formula", "name", "delta")

```

```{r message=FALSE, echo=FALSE, out.width="75%"}
library(kableExtra)
delta_AIC |>
  kbl() |>
  kable_classic_2("hover", full_width = F) |>
    row_spec(c(1), background = "lightblue")
```

By using `dredge` we find that there are a few candidate models that fit the data well.

The difference between the model AICs is lower than a common threshold of 2 but there is a difference between the top model and others. This model contains multiple interactions including `NDYI:Precip`, `NDYI:Temp`, `NDYI:WIP`	`Precip:Temp`, `Precip:WIP`, `Temp:WIP`,` NDYI:Precip:Temp`

We can visualize some of these interactions with interaction plots against SOC stock. It looks like there is strong interaction and separation with the `NDYI` parameter and `WIP`

```{r fig.show='hold', message=FALSE, echo=FALSE}
library(interactions)

precwip <- interactions::interact_plot(dmod1, pred = Precip, modx = WIP, plot.points = T, plot.shape = T,
                            x.label = "Precip", 
                            y.label = "Fitted Dredge Model SOC Stock",
                            legend.main = "WIP Range",
                            colors = "blue",
                            interval = TRUE, )
tempwip <- interactions::interact_plot(dmod1, pred = Temp, modx = WIP, plot.points = T, plot.shape = T,
                            x.label = "Temp", 
                            y.label = "Fitted Dredge Model SOC Stock",
                            legend.main = "WIP Range",
                            colors = "seagreen",
                            interval = TRUE, )
ndyiwip <- interactions::interact_plot(dmod1, pred = NDYI, modx = WIP, plot.points = T, plot.shape = T,
                            x.label = "NDYI", 
                            y.label = "Fitted Dredge Model SOC Stock",
                            legend.main = "WIP Range",
                            colors = "green",
                            interval = TRUE, )

ggarrange(precwip, tempwip,ndyiwip, ncol = 3, nrow=1, common.legend = T)

precndyi <- interactions::interact_plot(dmod1, pred = Precip, modx = NDYI, plot.points = T, plot.shape = T,
                            x.label = "Precip", 
                            y.label = "Fitted Dredge Model SOC Stock",
                            legend.main = "NDYI Range",
                            colors = "blue",
                            interval = TRUE )
tempndyi <- interactions::interact_plot(dmod1, pred = Temp, modx = NDYI, plot.points = T, plot.shape = T,
                            x.label = "Temp", 
                            y.label = "Fitted Dredge Model SOC Stock",
                            legend.main = "NDYI Range",
                            colors = "seagreen",
                            interval = TRUE )
wipndyi <- interactions::interact_plot(dmod1, pred = WIP, modx = NDYI, plot.points = T, plot.shape = T,
                            x.label = "WIP", 
                            y.label = "Fitted Dredge Model SOC Stock",
                            legend.main = "NDYI Range",
                            colors = "green",
                            interval = TRUE )
ggarrange(precndyi, tempndyi,wipndyi, ncol = 3, nrow=1, common.legend = T)


```

We can look at the bootstrapped confidence intervals to examine the significance of the predictors between the two models. Interestingly the `WIP` parameter alone is not significant. And there is high effect sizes for `NDYI`, `NDYI:Temp`, and `(NDYI × Precip) × Temp`

<dev align="center">
```{r cache=TRUE, echo=FALSE}
tab_model(dmod1, bootstrap = T, seed = 10, dv.labels = c("Dredge Model"))
```
</dev>


```{r include=FALSE}
r.sq(log10(wa_spl_dat_scale$SOC_stock_spline), fitted(dmod1))
```

The R$^{2}$ of the Dredge Model is R$^{2}$ = `r round(r.sq(log10(wa_spl_dat_scale$SOC_stock_spline), fitted(dmod1)), 3)` 

```{r echo=FALSE, message=FALSE}
dmod_wet <- lmer(log10(SOC_stock_spline) ~  WIP*Precip + NDYI+
                     GEOLOGIC_AGE+ lower_depth + (lower_depth||sample_ID), 
              data = wa_spl_dat_scale_wet, REML = F, na.action = 'na.fail')
dredge_wet <- dredge(dmod_wet, beta = "sd" )
dmodwet1 <- get.models(dredge_wet, 1)[[1]]

dmod_upl <- lmer(log10(SOC_stock_spline) ~  WIP*Precip+NDYI+
                     GEOLOGIC_AGE+ lower_depth + (lower_depth||sample_ID), 
              data = wa_spl_dat_scale_upl, REML = F, na.action = 'na.fail')
dredge_upl <- dredge(dmod_upl, beta = "sd" )
dmodupl1 <- get.models(dredge_upl, 1)[[1]]

tab_model(dmodwet1, dmodupl1, dv.labels = c("Wetlands (WIP > 0.50", "Uplands (WIP < 0.50)"))
```


### Random Forest Machine Learning


Using Random Forest can see if a more flexible, machine learning model can capture any non-linear relationships or interaction relationships between parameters.

`lower_depth` is included as a predictor here in the data setup

```{r message=FALSE, echo=FALSE}
library(randomForest)
set.seed(11)

# Validation Set 
train.index <- as.vector(sample(c(1:nrow(wa_spl_dat_scale)), 0.7*nrow(wa_spl_dat_scale), replace=F))
train <- wa_spl_dat_scale[train.index, c("SOC_stock_spline", "WIP", "GEOLOGIC_AGE", "Temp", "Precip", "lower_depth", "NDVI", "ANDWI", "EVI", "NDYI", "geomorphons" )]
train$GEOLOGIC_AGE <- as.character(train$GEOLOGIC_AGE)
train$SOC_stock_spline <- log10(train$SOC_stock_spline)
test <- wa_spl_dat_scale[-train.index, c("SOC_stock_spline", "WIP", "GEOLOGIC_AGE", "Temp", "Precip", "lower_depth", "NDVI", "ANDWI", "EVI", "NDYI", "geomorphons")]
test$GEOLOGIC_AGE <- as.character(test$GEOLOGIC_AGE)
test$SOC_stock_spline <- log10(test$SOC_stock_spline)
full <- wa_spl_dat_scale[, c("SOC_stock_spline", "WIP", "GEOLOGIC_AGE", "Temp", "Precip", "lower_depth", "NDVI", "ANDWI", "EVI", "NDYI",  "geomorphons")]
full$GEOLOGIC_AGE <- as.character(full$GEOLOGIC_AGE)
full$SOC_stock_spline <- log10(full$SOC_stock_spline)

```


We then use the `tuneRF` to choose the appropriate `mtry` number 

```{r echo=FALSE, message=FALSE, results='hide'}
set.seed(11)
tuneRF(x = full[, 2:ncol(full)], y = (full$SOC_stock_spline), mtryStart = 3, ntreeTry = 500, stepFactor = 1.5, improve = 0.0001)
```


```{r fig.show='hold', cache=TRUE, results='hide'}
rf_model <- randomForest((SOC_stock_spline) ~ .,
                         ntree = 1000, mtry = 4,
                         importance = TRUE, data = full)
plot(rf_model)

rf.full <- predict(rf_model, newdata = full)
vip::vip(rf_model)
```


61.63% of the out of bag (OOB) variation is explained. Looks like`Precip`, `lower_depth`, `WIP`, and `NDYI` are the big drivers. 


```{r echo=FALSE}

errdf <- data.frame("error variable" = c("MAE", "MAE^2", "RMSE", "RMSE non-log", "Stdev of non-log SOC stock",  "R^2 on full dataset"), 
                    "value" = c(caret::MAE(pred = rf.full, obs = (full$SOC_stock_spline)),
                                caret::MAE(pred = rf.full, obs = (full$SOC_stock_spline))**2,
                                caret::RMSE(pred = rf.full, obs = (full$SOC_stock_spline)),
                               exp(caret::RMSE(pred = rf.full, obs = (full$SOC_stock_spline))),
                                sd(wa_spl_dat_scale$SOC_stock_spline),
                                caret::R2(pred = rf.full, obs = (full$SOC_stock_spline))))
errdf |> kbl() |>
  kable_classic_2("hover", full_width = F)
```


Looking at partial dependency plots we can see how `Precip` and `NDYI` have higher influence on separating variation in model predictions. `WIP` less so but seems to play a role on the higher end. 


```{r tidy=FALSE, echo=FALSE, message=FALSE, cache=TRUE}
library(gridExtra)
library(grid)


wip_pdp<- data.frame(randomForest::partialPlot(rf_model, full, WIP, plot = FALSE)) |> 
    ggplot() +
        geom_line(aes(x = x, y = y)) +
    xlab("WIP") +
    ylim(-1, -0.25) +
    ylab(expression('Model Prediction Effect'))
precip_pdp<- data.frame(randomForest::partialPlot(rf_model, full, Precip, plot = FALSE)) |> 
    ggplot() +
        geom_line(aes(x = x, y = y))+
    xlab("Precipitation") +
    ylim(-1,  -0.25) +
    ylab(expression('Model Prediction Effect'))
ndyi_pdp <- data.frame(randomForest::partialPlot(rf_model, full, NDYI, plot = FALSE)) |> 
    ggplot() +
        geom_line(aes(x = x, y = y))+
    xlab("NDYI") +
    ylim(-1, -0.25) +
    ylab(expression('Model Prediction Effect'))
lower_pdp <- data.frame(randomForest::partialPlot(rf_model, full, lower_depth, plot = FALSE)) |> 
    ggplot() +
        geom_line(aes(x = x, y = y))+
    xlab("Depth") +
    ylim(-1,  -0.25) +
    ylab(expression('Model Prediction Effect'))



grid.arrange(precip_pdp, wip_pdp, lower_pdp, ndyi_pdp, ncol = 2, nrow=2)
```


```{r fig.show='hold', echo=FALSE}

rf.predtrain <- predict(rf_model, newdata = train)
rf.predtest <- predict(rf_model, newdata = test)
rf.predfull <- predict(rf_model, newdata = full)
```

```{r "RF R2", echo=FALSE, message=FALSE, include=FALSE, eval=FALSE}

# paste0(round(mean((rf.predtrain - (train$SOC_stock_spline))^2), 3), " mean square error rf model from training")
paste0(round(r.sq((train$SOC_stock_spline), rf.predtrain), 3), " R^2 rf model from training")
paste0(round(r.sq((test$SOC_stock_spline), rf.predtest), 3),  " R^2 rf model from testing ")

```



```{r fig.show='hold', cache=TRUE, results='hide'}
fullwet <- wa_spl_dat_scale_wet[, c("SOC_stock_spline", "WIP", "GEOLOGIC_AGE", "Temp", "Precip", "lower_depth", "NDVI", "ANDWI", "EVI", "NDYI",  "geomorphons")]
fullwet$GEOLOGIC_AGE <- as.character(fullwet$GEOLOGIC_AGE)
fullwet$SOC_stock_spline <- log10(fullwet$SOC_stock_spline)

fullupl <- wa_spl_dat_scale_upl[, c("SOC_stock_spline", "WIP", "GEOLOGIC_AGE", "Temp", "Precip", "lower_depth", "NDVI", "ANDWI", "EVI", "NDYI",  "geomorphons")]
fullupl$GEOLOGIC_AGE <- as.character(fullupl$GEOLOGIC_AGE)
fullupl$SOC_stock_spline <- log10(fullupl$SOC_stock_spline)

rf_modelwet <- randomForest((SOC_stock_spline) ~ .,
                         ntree = 1000, mtry = 4,
                         importance = TRUE, data = fullwet)
plot(rf_modelwet)

rf.fullwet <- predict(rf_modelwet, newdata = fullwet)
wetvip <- vip::vip(rf_modelwet)
```
```{r}
rf_modelupl <- randomForest((SOC_stock_spline) ~ .,
                         ntree = 1000, mtry = 4,
                         importance = TRUE, data = fullupl)
plot(rf_modelupl)

rf.fullupl <- predict(rf_modelupl, newdata = fullupl)
uplvip <- vip::vip(rf_modelupl)
```


```{r}
vipwetupl <- ggarrange(wetvip, uplvip, ncol = 2, labels = c("Wetland", "Upland") )

ggplot2::ggsave(plot = vipwetupl, filename = paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/vipwetupl.jpeg"),
                width = 9, height = 7.5, units = "in", dpi = 500)
```



We can now visualize our candidate model 2.1 in a predicted vs. actual plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(RColorBrewer)
mod2.1graph <- ggplot(wa_spl_dat_scale, 
                      aes(y = 10**(fitted(mod2.1)), x = (SOC_stock_spline))) +
    geom_jitter(color='black', 
                aes(fill = (wa_spl_dat_0_30_60_100$WIP*100), 
                    shape = as.factor(GEOLOGIC_AGE)),
               size = 3, stroke = 0.9, alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlGnBu"),
                         name = "WIP %", n.breaks = 5, limits = c(0, 100)) +
     scale_shape_manual(name = "Geologic Age", 
                        values = c(20, 21, 22, 23, 24, 25),
                        labels= c("Pleistocene", "Quaternary", 
                                  "Miocene-Eocene", "Oligocene-Eocene",
                                  "Eocene", "pre-Tertiary" )) +
    geom_smooth(aes(y = 10**(fitted(mod2.1)), x = (SOC_stock_spline)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) +
    xlab(expression('Sampled SOC Stock (g cm'^-2*')')) + 
    ylab(expression('Predicted SOC Stock (g cm'^-2*')')) + 
    geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed") +
    annotate("text", 
             label = paste("R^{2} == ", 
                           round(r.sq((wa_spl_dat_scale$SOC_stock_spline),
                                      10**fitted(mod2.2)), 3)), 
             x = 0.5, y = 2.2, size = 4, parse = T) + 
    #xlim(-4, 5) +
    #ylim(-4, 5)  +
    guides(guide_legend(byrow = TRUE)) +
    theme(legend.position = 'right', 
          legend.key.size = unit(0.5, "cm"),
          panel.background = element_blank(),
          panel.grid.major = element_line(colour = "grey80"),
          axis.ticks = element_blank(),
          text = element_text(size = 18))

ggplot2::ggsave(plot = mod2.1graph, filename = paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/model2_1.png"),
                width = 9, height = 7.5, units = "in", dpi = 500)
```


We now can examine the plot with the Dredge Model
```{r tidy=FALSE, echo=FALSE, message=FALSE, include=FALSE}

dmodgraph <- ggplot(wa_spl_dat_scale, 
                    aes(y = 10**(fitted(dmod1)), x = (SOC_stock_spline))) +
    geom_jitter(color='black', 
               aes(fill = (wa_spl_dat_0_30_60_100$WIP*100), 
                   shape = as.factor(GEOLOGIC_AGE)),
               size = 3, stroke = 0.9, alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlGnBu"), 
                         name = "WIP %", n.breaks = 5, limits = c(0, 100)) +
     scale_shape_manual(name = "Geologic Age", 
                        values = c(20, 21, 22, 23, 24, 25),
                        labels= c("Pleistocene", "Quaternary", 
                                  "Miocene-Eocene", "Oligocene-Eocene",
                                  "Eocene", "pre-Tertiary" )) +
    geom_smooth(aes(y = 10**(fitted(dmod1)), x = (SOC_stock_spline)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) +
    xlab(expression('Sampled SOC Stock (g cm'^-2*')')) + 
    ylab(expression('Predicted SOC Stock (g cm'^-2*')')) + 
    geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed") +
    annotate("text", label = paste("R^{2} == ", 
                                   round(r.sq((wa_spl_dat_scale$SOC_stock_spline), 10**fitted(dmod1)), 3)), 
             y = 2.2, x = 0.5, size = 4, parse = T) + 
    guides(guide_legend(byrow = TRUE)) +
    theme(legend.position = 'right', 
          legend.key.size = unit(0.5, "cm"),
          panel.background = element_blank(),
          panel.grid.major = element_line(colour = "grey80"),
          axis.ticks = element_blank(),
          text = element_text(size = 18))

ggplot2::ggsave(plot = dmodgraph, filename = paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/dmod.png"),
                width = 9, height = 7.5, units = "in", dpi = 500)
```



Here, we show the fit between sampled and predicted SOC stocks for the full Random Forest Model

```{r fig.show='hold', tidy=FALSE, echo=FALSE, message=FALSE, include=FALSE}

library(RColorBrewer)
rf_graph <- ggplot(wa_spl_dat_scale, 
                   aes(y = 10**(rf.full), x = (SOC_stock_spline))) +
    geom_jitter(color='black', 
               aes(fill = (wa_spl_dat_0_30_60_100$WIP*100), 
                   shape = as.factor(GEOLOGIC_AGE)),
               size = 3, stroke = 0.9, alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlGnBu"), 
                         name = "WIP %", n.breaks = 5, limits = c(0, 100)) +
    scale_shape_manual(name = "Geologic Age", 
                        values = c(20, 21, 22, 23, 24, 25),
                        labels= c("Pleistocene", "Quaternary", 
                                  "Miocene-Eocene", "Oligocene-Eocene",
                                  "Eocene", "pre-Tertiary" )) +
    geom_smooth(aes(y = 10**(rf.full), x = (SOC_stock_spline)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) +
    xlab(expression('Sampled SOC Stock (g cm'^-2*')')) + 
    ylab(expression('Predicted SOC Stock (g cm'^-2*')')) + 
    geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed") +
    annotate("text", label = paste("R^{2} == ",
                                   round(r.sq((wa_spl_dat_scale$SOC_stock_spline), 10**(rf.predfull)), 3)),
             y = 2.2, x = 0.5, size = 4, parse = T) +
    #xlim(-4, 5) +
    #ylim(-4, 5)  +
    guides(guide_legend(byrow = TRUE)) +
    theme(legend.position = 'right', 
          legend.key.size = unit(0.5, "cm"),
          panel.background = element_blank(),
          panel.grid.major = element_line(colour = "grey80"),
          axis.ticks = element_blank(),
          text = element_text(size = 18))

ggplot2::ggsave(plot = rf_graph, filename = paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/rf_model.png"),
                width = 9, height = 7.5, units = "in", dpi = 500)
```

```{r fig.show='hold', tidy=FALSE, echo=FALSE, message=FALSE, include=FALSE, out.height=12}

allpredgraph <- ggpubr::ggarrange(mod2.1graph, dmodgraph, rf_graph, common.legend = T, ncol = 1, nrow= 3)

ggplot2::ggsave(plot = allpredgraph, filename = paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/allpred_graph.png"),
                width = 9, height = 18, units = "in", dpi = 500)
```


```{r}
library(ppcor)

pcor.test(wa_spl_dat_scale$WIP, wa_spl_dat_scale$SOC_stock_spline, wa_spl_dat_scale$lower_depth)

pcor.test(wa_spl_dat_scale$Precip, wa_spl_dat_scale$SOC_stock_spline, wa_spl_dat_scale$lower_depth)

pcor.test(wa_spl_dat_scale$Temp, wa_spl_dat_scale$SOC_stock_spline, wa_spl_dat_scale$lower_depth)

pcor.test(wa_spl_dat_scale$EVI, wa_spl_dat_scale$SOC_stock_spline, wa_spl_dat_scale$lower_depth)
```


