---
title: "SOC controls across Washington"
author: "Anthony Stewart"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}
```

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.show = "hold", time_it = TRUE, dpi = 75)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE, collapse = TRUE)
knitr::opts_knit$set(root.dir = '/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/')
library(rgl)
library(terra)
library(lme4)
library(MASS)
library(mgcv)
library(lmerTest)
library(MuMIn)
library(RLRsim)
library(terra)
library(spatialEco)
library(sf)
library(mapview)
library(tidyr)
library(car)
library(ggplot2)
library(sjPlot)
library(sjstats)
library(DHARMa)
library(ggeffects)
library(merTools)
library(glmnet)
library(stats)
library(ggcorrplot)
library(RColorBrewer)
library(cowplot)
library(webshot)
library(kableExtra)
library(caret)
library(pdp)
library(vip)
library(formatR)
library(dplyr)

knitr::knit_hooks$set(webgl = hook_webgl)
rgl::setupKnitr(autoprint = TRUE)
```

```{r include=FALSE}

# Quick R^2 function
r.sq <- function(y,y.fitted){
    res <- y-y.fitted
    1-sum(res^2)/sum((y-mean(y))^2)
}
```

### Introduction

Mapping soil organic carbon (SOC) as a function climatic, topographic, and geologic factors across spatial extents is key for identifying vulnerable SOC stocks. Yet, large, vulnerable SOC stocks are contained in the organic soils of wetlands and can occur in close proximity to non-wetland soils but are controlled and function differently due to saturated soil conditions. Thus, wetland SOC is often omitted or underrepresented in SOC mapping and modeling making it difficult to assess controls compared to non-wetland soil types. Here, we use geospatial data and SOC stocks from 481 soil horizons in 96 pedons in both wetlands and non-wetlands across three ecoregion study areas of the Pacific Northwest to assess climatic, geologic, and topographic controls on SOC. 

### Hypotheses

1.  WIP is a significant predictor of SOC stock

2.  WIP interaction with climate (Precip & Temp) is a significant predictor of SOC stock

3.  Geology is a significant predictor of SOC stock

### Data
This dataset has observations of SOC% and SOC stock for 96 locations in Washington state. There are geospatial covariates as well that are used to build models to predict SOC% and/or SOC stock


```{r echo=FALSE}
wa_dat <- read.csv("SOIL CARBON/ANALYSIS/All_WA_horizons_spec_geoage.csv")
str(wa_dat)
```

Examining covariate predictors using correlation plots

```{r echo=FALSE}

wa_dat_num <- wa_dat |> dplyr::select(carbon_perc,
                               carbon_stock_g_cm2,
                               depth_cm,
                               Precip, 
                               Temp,
                               NDYI,
                               NDVI,
                               MNDWI,
                               EVI,
                               SCI,
                               WIP) |> as.matrix()

ggcorrplot(cor(wa_dat_num), method = "square", type = "full", lab = T, lab_size = 2)

```

There is weak correlation between `carbon_stock_g_cm2` and the selected covariates. Additionally there is some collinearity between predictors. The spectral predictors such as EVI (Enhanced Vegetation Index) and SCI (Soil Condition Index) are well correlated with each other and also with NDVI (Normalized Difference Vegetation Index). We can remove these based on the correlation coefficient > 0.7.

Now we can begin examining the distribution of carbon stock values in the dataset to choose the appropriate transformation. We also scale and center the numeric predictor variables.

```{r fig.align='center', echo=FALSE}
columns_to_exclude <- c("carbon_perc", "carbon_stock_g_cm2") 

wa_dat_scale <- wa_dat |> 
    dplyr::mutate(across(
        dplyr::where(is.numeric) & !all_of(columns_to_exclude),
                  ~dplyr::case_when(TRUE ~ scale(.))))
str(wa_dat_scale)
hist(wa_dat_scale$carbon_perc) #not of interest but taking a look 
hist(wa_dat_scale$carbon_stock_g_cm2)

```

### Explicit parameter model building 


Now build models using log transformed carbon stock data. We need to specify that `sample_ID` is a random effect because of the multiple samples at one location. `depth` can be a random slope to adjust model based on how it is affected by depth. We could use a Generalized Linear Model or a Generalized Linear Mixed Model here too but they often fail to converge. I am starting with the log-transformed linear mixed model first to test the hypotheses

```{r}
#Full model with hypothesized parameters
mod1 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP*Precip*Temp +  MNDWI + NDVI + NDYI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
# No interactions
mod2 <-lmer(log(carbon_stock_g_cm2) ~ 
                WIP+Precip+Temp + MNDWI + NDVI + NDYI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No spectral
mod3 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP*Precip*Temp + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No geology
mod4 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP*Precip*Temp +  MNDWI + NDVI + NDYI + 
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No WIP w/ interaction
mod5 <- lmer(log(carbon_stock_g_cm2) ~ 
                 Precip*Temp +  MNDWI + NDVI + NDYI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No WIP w/o interaction
mod6 <- lmer(log(carbon_stock_g_cm2) ~ 
                 Precip+Temp + MNDWI + NDVI + NDYI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No climate
mod7 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP+  MNDWI + NDVI + NDYI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No climate No spectral
mod8 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#Just WIP
mod9 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP + 
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)

```

Pairwise comparisons between the top, global model and the rest. Note the table contains mod1 multiple times and has been encoded with an additional number 

```{r tidy=FALSE, echo=FALSE}
anova_table_mod1 <- rbind(anova(mod1, mod2), #Not significant
                            anova(mod1, mod3), #
                            anova(mod1, mod4), #Significant
                            anova(mod1, mod5),#Significant
                            anova(mod1, mod6), #Significant
                            anova(mod1, mod7), #Significant
                            anova(mod1, mod8), #Significant
                            anova(mod1, mod9)) #Significant

anova_table_mod1 |>
    tibble::rownames_to_column(var = "models") |> 
    dplyr::mutate(Significant = case_when(`Pr(>Chisq)` < 0.05 & `Pr(>Chisq)`>  0.01~ "*",
                                `Pr(>Chisq)` < 0.01 & `Pr(>Chisq)` > 0.001~ "**",
                                `Pr(>Chisq)` < 0.001 ~ "***",
                                .default = ""),
           Models = case_when(models == "mod11" ~ "mod1",
                              models == "mod12" ~ "mod1",
                              models == "mod13" ~ "mod1",
                              models == "mod14" ~ "mod1",
                              models == "mod15" ~ "mod1",
                              models == "mod16" ~ "mod1",
                              models == "mod17" ~ "mod1",
                              .default = models)) |> 
    dplyr::select(-models) |> 
    dplyr::select(Models, everything()) |> 
    kbl() |>
    kable_classic_2("hover", full_width = F) 
```
From the ANOVAs it looks like `model 2` & `model 3` performed just as well as `model 1`meaning we should proceed with these. `model 2` removed interactions and `model 3` removed all the spectral metrics but kept interaction. However, `model 2` has 1 less parameter than `model 3` which gives it a lower AIC when compared together. 

```{r echo=FALSE}
anova(mod3, mod2) |> 
    kbl() |>
    kable_classic_2("hover", full_width = F) 
```


```{r}
#No MNDWI
mod2.1 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP+Precip+Temp + NDVI + NDYI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No MNDWI, No NDVI
mod2.2 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP+Precip+Temp + NDYI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No NDYI, No NDVI
mod2.3 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP+Precip+Temp + MNDWI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No NDYI, No MNDWI
mod2.4 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP+Precip+Temp + NDVI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No NDVI
mod2.5 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP+Precip+Temp + NDYI + MNDWI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No NDYI
mod2.6 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP+Precip+Temp + NDVI + MNDWI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No spectral
mod2.7 <- lmer(log(carbon_stock_g_cm2) ~ 
                 WIP+Precip+Temp + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
```

```{r tidy=FALSE, echo=FALSE}
anova_table_mod2 <- rbind(anova(mod2, mod2.1), #Not Significant
anova(mod2, mod2.2), #Not Significant
anova(mod2, mod2.3), #Not Significant
anova(mod2, mod2.4), #Not Significant
anova(mod2, mod2.5), #Not Significant
anova(mod2, mod2.6), 
anova(mod2, mod2.7), make.row.names = T) #Not Significant

anova_table_mod2 |>
    tibble::rownames_to_column(var = "models") |> 
    mutate(Significant = case_when(`Pr(>Chisq)` < 0.05 & `Pr(>Chisq)`>  0.01~ "*",
                                `Pr(>Chisq)` < 0.01 & `Pr(>Chisq)` > 0.001~ "**",
                                `Pr(>Chisq)` < 0.001 ~ "***",
                                .default = ""),
           Models = case_when(models == "mod21" ~ "mod2",
                              models == "mod22" ~ "mod2",
                              models == "mod23" ~ "mod2",
                              models == "mod24" ~ "mod2",
                              models == "mod25" ~ "mod2",
                              models == "mod26" ~ "mod2",
                              .default = models)) |>
    dplyr::select(-models) |> 
    dplyr::select(Models, everything()) |> 
    kbl() |>
    kable_classic_2("hover", full_width = F)

```
Model 2.2 and 2.3 are the lowest AIC and not significantly different from the fit in Model 2. The choice is either between `MNDWI` or `NDYI` to be included in the model with `WIP`, `Precip`, `Temp`, and `Geology.` To test , I want to see if `WIP` interacts with these variables. 

```{r}
#No MNDWI, No NDVI
mod2.2.1 <- lmer(log(carbon_stock_g_cm2) ~ 
                 Precip+Temp + WIP:NDYI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
#No NDYI, No NDVI
mod2.3.1 <- lmer(log(carbon_stock_g_cm2) ~ 
                 Precip+Temp + WIP:MNDWI + (GEOLOGIC_AGE)+
                 (1+(depth_cm)|sample_ID), data = wa_dat_scale, REML = F)
```

```{r echo=FALSE}
anova_table_mod2ext <- rbind(anova(mod2.2, mod2.2.1),
anova(mod2.3, mod2.3.1))

anova_table_mod2ext |>
    mutate(Significant = case_when(`Pr(>Chisq)` < 0.05 & `Pr(>Chisq)`>  0.01~ "*",
                                `Pr(>Chisq)` < 0.01 & `Pr(>Chisq)` > 0.001~ "**",
                                `Pr(>Chisq)` < 0.001 ~ "***",
                                .default = "")) |> 
  kbl() |>
  kable_classic_2("hover", full_width = F)
```

Doesn't look like there are any significant interactions that improve model fit

Now we can look at the table of all models and compare AICs

```{r 'model list and table', message=FALSE, echo=FALSE}
mod_list <- list(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod2.1, mod2.2,
                 mod2.3, mod2.4, mod2.5, mod2.6, mod2.7, mod2.2.1, mod2.3.1)
mod_names <- c("mod1", "mod2", "mod3", "mod4", "mod5", "mod6", "mod7", "mod8", "mod9",
               "mod2.1", "mod2.2", "mod2.3", "mod2.4", "mod2.5", "mod2.6", "mod2.7", "mod2.2.1", "mod2.3.1")


mod_AIC = matrix(nrow = length(mod_list), ncol =3)


for (i in 1:length(mod_list)) {
    mod_AIC[i,1] <- round(AIC(mod_list[[i]]), 2)
    mod_AIC[i,2] <- paste0(formula(mod_list[[i]]))[3]
    mod_AIC[i,3] <- mod_names[i]
}
    
delta_AIC <- as.data.frame(cbind(mod_AIC))
delta_AIC$delta <- as.numeric(delta_AIC$V1) -  min(as.numeric(delta_AIC$V1))
names(delta_AIC) <- c("AIC", "formula", "name", "delta")



#tab_model(mod_list, show.aic = T, show.aicc = T, dv.labels = mod_names)
```

```{r message=FALSE, echo=FALSE, out.width="75%"}
library(kableExtra)
delta_AIC |>
  kbl() |>
  kable_classic_2("hover", full_width = F) |>
    row_spec(c(11, 12), background = "lightblue")
```

Looks like the best model fit to the data according to AIC is the `mod2.2` or `mod2.3` which do not include any interactions. The model $R^2$ for `mod2.2` is `r round(r.sq(log(wa_dat$carbon_stock_g_cm2), fitted(mod2.2)), 3)` compared to the $R^2$ for `mod2.3` is `r round(r.sq(log(wa_dat$carbon_stock_g_cm2), fitted(mod2.3)),3)`

```{r include=FALSE}
cor(wa_dat_scale$MNDWI, wa_dat_scale$NDYI)
round(r.sq(log(wa_dat_scale$carbon_stock_g_cm2), fitted(mod2.2)), 3)
round(r.sq(log(wa_dat_scale$carbon_stock_g_cm2), fitted(mod2.3)), 3)
```


Confidence intervals calculated with bootstrapping between these two models show that `NDYI` is a significant predictor compared to `MNDWI`. Therefore, we use `mod2.2` as our candidate model


```{r cache=TRUE, echo=FALSE, out.width="75%"}
tab_model(mod2.2, mod2.3, bootstrap = T, seed = 10, dv.labels = c("Model 2.2", "Model 2.3"))
```

--- 

We can now visualize our candidate model 2.2. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(RColorBrewer)
ggplot(wa_dat_scale, aes(y = (fitted(mod2.2)), x = log(carbon_stock_g_cm2))) +
    geom_jitter(color='black', 
               aes(fill = (wa_dat$WIP*100), shape = as.factor(GEOLOGIC_AGE)),
               size = 3, stroke = 0.9, alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlGnBu"),
                         name = "WIP %", n.breaks = 5, limits = c(0, 100)) +
     scale_shape_manual(name = "Geologic Age", 
                        values = c(20, 21, 22, 23, 24, 25),
                        labels= c("Pleistocene", "Quaternary", 
                                  "Miocene-Eocene", "Oligocene-Eocene",
                                  "Eocene", "pre-Tertiary" )) +
    geom_smooth(aes(y = (fitted(mod2.2)), x = log(carbon_stock_g_cm2)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) +
    xlab(expression('Log Sampled SOC Stock (g cm'^-2*')')) + 
    ylab(expression('Log Predicted SOC Stock (g cm'^-2*')')) + 
    geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed") +
    annotate("text", label = paste("R^{2} == ", 
                                   round(r.sq(log(wa_dat_scale$carbon_stock_g_cm2), fitted(mod2.2)), 3)), 
             x = -5, y = 1, size = 4, parse = T) + 
    #xlim(-4, 5) +
    #ylim(-4, 5)  +
    guides(guide_legend(byrow = TRUE)) +
    theme(legend.position = 'right', 
          legend.key.size = unit(0.5, "cm"),
          panel.background = element_blank(),
          panel.grid.major = element_line(colour = "grey80"),
          axis.ticks = element_blank(),
          text = element_text(size = 18))

ggplot2::ggsave( paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/model2_2.png"),
                width = 9, height = 7.5, units = "in", dpi = 500)
```

### Dredging for model selection

The `dredge` function can be used to look through multiple combinations of models from a globally defined model. The number of parameters are limited so we run two models: one with the `NDYI` parameter and another with the `MNDWI` parameter included with `Temp`, `Precip`, and `WIP` fully interacting. `GEOLOGIC_AGE` is also included as an additional, non-interaction parameter and random effects are constant.

```{r, cache=TRUE, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
gmod1 <- lmer(log(carbon_stock_g_cm2) ~  WIP*Temp*Precip*NDYI+GEOLOGIC_AGE+ (1+(depth_cm)|sample_ID), 
              data = wa_dat_scale, REML = F, na.action = 'na.fail')
gmod2 <- lmer(log(carbon_stock_g_cm2) ~  WIP*Temp*Precip*MNDWI+GEOLOGIC_AGE+ (1+(depth_cm)|sample_ID), 
              data = wa_dat_scale, REML = F, na.action = 'na.fail')
dredge1 <- dredge(gmod1, beta = "sd" )
dredge2 <- dredge(gmod2, beta = "sd" )
```

```{r, cache=TRUE, out.height="50%", warning=FALSE, echo=FALSE, out.width="75%"}
head(dredge1) |>
  kbl() |>
  kable_classic_2("hover", full_width = F) |>
    scroll_box(width = "100%")
dmod1 <- get.models(dredge1, 1)[[1]]

head(dredge2) |>
  kbl() |>
  kable_classic_2("hover", full_width = F)|>
    scroll_box(width = "100%")
dmod2 <- get.models(dredge2, 1)[[1]]
```

By using `dredge` we find that there are a few candidate models that fit the data well.

The difference between the model AICs is negligible. The first includes no interactions between the parameters and removes `Temp` as a predictor. The second includes an interaction between `MNDWI` and `Precip` while also excluding `Temp`. 

We can look at the bootstrapped confidence intervals to examine the significance of the predictors between the two models. 
```{r cache=TRUE, echo=FALSE}
tab_model(dmod1, dmod2, bootstrap = T, seed = 10, dv.labels = c("Dredge Model 1", "Dredge Model 2"))
```

Dredge Model 1 includes `NDYI`, `Precip`, `WIP`, and `GEOLOGIC_AGE` as significant predictors. Dredge Model 2 does not have significant confidence intervals for the `MNDWI` and `MNDWIxPrecip` parameters. 

```{r include=FALSE}
r.sq(log(wa_dat_scale$carbon_stock_g_cm2), fitted(dmod1))
r.sq(log(wa_dat_scale$carbon_stock_g_cm2), fitted(dmod2))
```

There is also minimal differences between the R$^{2}$ between the two models where Dredge Model 1 has R$^{2}$ = `r round(r.sq(log(wa_dat$carbon_stock_g_cm2), fitted(dmod1)), 3)` and Dredge Model 2 has R$^{2}$ = `r round(r.sq(log(wa_dat$carbon_stock_g_cm2), fitted(dmod2)), 3)`

We now can examine the plot with the Dredge Model 1
```{r tidy=FALSE, echo=FALSE, message=FALSE}

ggplot(wa_dat_scale, aes(y = (fitted(dmod1)), x = log(carbon_stock_g_cm2))) +
    geom_jitter(color='black', 
               aes(fill = (wa_dat$WIP*100), shape = as.factor(GEOLOGIC_AGE)),
               size = 3, stroke = 0.9, alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlGnBu"), 
                         name = "WIP %", n.breaks = 5, limits = c(0, 100)) +
     scale_shape_manual(name = "Geologic Age", 
                        values = c(20, 21, 22, 23, 24, 25),
                        labels= c("Pleistocene", "Quaternary", 
                                  "Miocene-Eocene", "Oligocene-Eocene",
                                  "Eocene", "pre-Tertiary" )) +
    geom_smooth(aes(y = (fitted(dmod1)), x = log(carbon_stock_g_cm2)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) +
    xlab(expression('Log Sampled SOC Stock (g cm'^-2*')')) + 
    ylab(expression('Log Predicted SOC Stock (g cm'^-2*')')) + 
    geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed") +
    annotate("text", label = paste("R^{2} == ", 
                                   round(r.sq(log(wa_dat_scale$carbon_stock_g_cm2), fitted(dmod1)), 3)), 
             x = -5, y = 1, size = 4, parse = T) + 
    guides(guide_legend(byrow = TRUE)) +
    theme(legend.position = 'right', 
          legend.key.size = unit(0.5, "cm"),
          panel.background = element_blank(),
          panel.grid.major = element_line(colour = "grey80"),
          axis.ticks = element_blank(),
          text = element_text(size = 18))

ggplot2::ggsave( paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/dmod.png"),
                width = 9, height = 7.5, units = "in", dpi = 500)
```


### Random Forest Machine Learning


I also want to try Random Forest in order to see if a more flexible, machine learning model can capture any non-linear relationships with SOC and other variables.

`depth_cm` is included as a predictor here in the data setup

```{r message=FALSE, echo=FALSE}
library(randomForest)
set.seed(11)

# Validation Set 
train.index <- as.vector(sample(c(1:nrow(wa_dat_scale)), 0.7*nrow(wa_dat_scale), replace=F))
train <- wa_dat_scale[train.index, c("carbon_stock_g_cm2", "WIP", "GEOLOGIC_AGE", "Temp", "Precip", "depth_cm", "NDVI", "MNDWI", "EVI", "NDYI", "Geomorphon_Class" )]
train$GEOLOGIC_AGE <- as.character(train$GEOLOGIC_AGE)
train$carbon_stock_g_cm2 <- log(train$carbon_stock_g_cm2)
test <- wa_dat_scale[-train.index, c("carbon_stock_g_cm2", "WIP", "GEOLOGIC_AGE", "Temp", "Precip", "depth_cm", "NDVI", "MNDWI", "EVI", "NDYI", "Geomorphon_Class")]
test$GEOLOGIC_AGE <- as.character(test$GEOLOGIC_AGE)
test$carbon_stock_g_cm2 <- log(test$carbon_stock_g_cm2)
full <- wa_dat[, c("carbon_stock_g_cm2", "WIP", "GEOLOGIC_AGE", "Temp", "Precip", "depth_cm", "NDVI", "MNDWI", "EVI", "NDYI",  "Geomorphon_Class")]
full$GEOLOGIC_AGE <- as.character(full$GEOLOGIC_AGE)
full$carbon_stock_g_cm2 <- log(full$carbon_stock_g_cm2)

```


We then use the `tuneRF` to choose the appropriate `mtry` number 

```{r echo=FALSE, message=FALSE, results='hide'}
set.seed(11)
tuneRF(x = full[, 2:ncol(full)], y = (full$carbon_stock_g_cm2), mtryStart = 3, ntreeTry = 500, stepFactor = 1.5, improve = 0.0001)
```


```{r fig.show='hold', cache=TRUE, results='hide'}
rf_model <- randomForest((carbon_stock_g_cm2) ~ .,
                         ntree = 1000, mtry = 6,
                         importance = TRUE, data = full)
plot(rf_model)

rf.full <- predict(rf_model, newdata = full)
vip::vip(rf_model)
```


~50% of the out of bag (OOB) variation is explained. Looks like after `depth_cm` `Precip`, `Temp`, `WIP`, and `NDYI` are the big drivers. The RF model does not appear to have the greatest fit... 

# EDIT HERE


```{r echo=FALSE}

errdf <- data.frame("error variable" = c("MAE", "MAE^2", "RMSE", "RMSE non-log", "Stdev of non-log SOC stock",  "R^2 on full dataset"), 
                    "value" = c(caret::MAE(pred = rf.full, obs = full$carbon_stock_g_cm2),
                                caret::MAE(pred = rf.full, obs = full$carbon_stock_g_cm2)**2,
                                caret::RMSE(pred = rf.full, obs = full$carbon_stock_g_cm2),
                                exp(caret::RMSE(pred = rf.full, obs = full$carbon_stock_g_cm2)),
                                sd(wa_dat_scale$carbon_stock_g_cm2),
                                caret::R2(pred = rf.full, obs = full$carbon_stock_g_cm2)))
errdf |> kbl() |>
  kable_classic_2("hover", full_width = F)
```


Look at partial dependency plots 

# EDIT HERE

```{r tidy=FALSE, echo=FALSE, message=FALSE, cache=TRUE}
library(gridExtra)
library(grid)
#randomForest::partialPlot(rf_model, full, WIP)
# randomForest::partialPlot(rf_model, full, Precip)
# randomForest::partialPlot(rf_model, full, Temp)
# randomForest::partialPlot(rf_model, full, NDYI)
# randomForest::partialPlot(rf_model, full, MNDWI)

precip_pdp <- rf_model |> pdp::partial(pred.var = c("Precip"), train = full) |> 
    ggplot2::autoplot(rug = T, train = full) +
    #xlab(expression('Sampled SOC %'))+ 
    ylim(-3, -1) +
    ylab(expression('Model Prediction Effect'))
wip_pdp <- rf_model |> pdp::partial(pred.var = c("WIP"), train = full) |> 
    ggplot2::autoplot(rug = T, train = full) +
    ylim(-3, -1) +
    #xlab(expression('Sampled SOC %'))+ 
    ylab(expression('Model Prediction Effect'))
evi_pdp <- rf_model |> pdp::partial(pred.var = c("EVI"), train = full) |> 
    ggplot2::autoplot(rug = T, train = full) +
    ylim(-3, -1) +
    #xlab(expression('Sampled SOC %'))+ 
    ylab(expression('Model Prediction Effect'))
ndyi_pdp <- rf_model |> pdp::partial(pred.var = c("NDYI"), train = full) |> 
    ggplot2::autoplot(rug = T, train = full) +
    ylim(-3, -1) +
    #xlab(expression('Sampled SOC %'))+ 
    ylab(expression('Model Prediction Effect'))

grid.arrange(precip_pdp, wip_pdp, evi_pdp, ndyi_pdp, ncol = 2, nrow=2)
```


```{r fig.show='hold', echo=FALSE}
rf.predtrain <- predict(rf_model, newdata = train)
rf.predtest <- predict(rf_model, newdata = test)
```

```{r "RF R2", echo=FALSE, message=FALSE}

paste0(round(mean((rf.predtrain - (train$carbon_stock_g_cm2))^2), 3), " mean square error rf model from training")
paste0(round(r.sq((train$carbon_stock_g_cm2), rf.predtrain), 3), " R^2 rf model from training")
paste0(round(r.sq((test$carbon_stock_g_cm2), rf.predtest), 3),  " R^2 rf model from testing ")

```

Here, we show the fit between sampled and predicted SOC stocks for the full Random Forest Model

```{r fig.show='hold', tidy=FALSE, echo=FALSE}

library(RColorBrewer)
ggplot(wa_dat_scale, aes(y = (rf.full), x = log(carbon_stock_g_cm2))) +
    geom_jitter(color='black', 
               aes(fill = (wa_dat$WIP*100), shape = as.factor(GEOLOGIC_AGE)),
               size = 3, stroke = 0.9, alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlGnBu"), 
                         name = "WIP %", n.breaks = 5, limits = c(0, 100)) +
    scale_shape_manual(name = "Geologic Age", 
                        values = c(20, 21, 22, 23, 24, 25),
                        labels= c("Pleistocene", "Quaternary", 
                                  "Miocene-Eocene", "Oligocene-Eocene",
                                  "Eocene", "pre-Tertiary" )) +
    geom_smooth(aes(y = (rf.full), x = log(carbon_stock_g_cm2)), 
                method = "lm", color = "#fa3e3e", fill = "#fa3e3e", 
                size = 1.2, linetype = 5, alpha = 0.3, se = T) +
    xlab(expression('Log Sampled SOC Stock (g cm'^-2*')')) + 
    ylab(expression('Log Predicted SOC Stock (g cm'^-2*')')) + 
    geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed") +
    #xlim(-4, 5) +
    #ylim(-4, 5)  +
    guides(guide_legend(byrow = TRUE)) +
    theme(legend.position = 'right', 
          legend.key.size = unit(0.5, "cm"),
          panel.background = element_blank(),
          panel.grid.major = element_line(colour = "grey80"),
          axis.ticks = element_blank(),
          text = element_text(size = 18))

ggplot2::ggsave( paste0("/Users/Anthony/OneDrive - UW/University of Washington/Presentations/Conferences/AGU2023/ExploringControls/Figures and Tables/rf_model.png"),
                width = 9, height = 7.5, units = "in", dpi = 500)
```
