---
title: "Creating Wetland Probability from ArcGIS outputs"
format: html
---

```{r setup, include=FALSE}
#| label: setup

set.seed(11)
library(formatR)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.show = "hold", time_it = TRUE, dpi = 75)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = T, collapse = TRUE)
knitr::opts_knit$set(root.dir = '/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/')
library(terra)
library(tidyverse)
library(tidyterra)
library(randomForest)
library(caret)
library(leaflet)
```

Options for speed
```{r}
terraOptions(
    memfrac = 0.1
)
```


## Mashel Wetland Probability

#### Data

Rasters 
```{r eval=FALSE}
filepath <- "Mashel/Mashel_WIP_2024/DataExport"
mas_dtw <- "Mashel/Mashel_WIP_2024/DataExport/Mashel_DTW.tif"

tif_proc <- function(filepath, target_rast){
    filelist <- list.files(filepath, pattern = ".tif$", full.names = TRUE) |> 
    str_subset(pattern = "^(?!.*flow)") |>
    str_subset(pattern = "^(?!.*fill)")  |>
    str_subset(pattern = "^(?!.*slopetwi.tif)") |>
    str_subset(pattern = "^(?!.*HUC12.tif)") |>
    str_subset(pattern = "^(?!.*filled)") |> 
    str_subset(pattern = "^(?!.*huc.tif)") |>
    str_subset(pattern = "^(?!.*Mosaic)")
    
    for(i in 1:length(filelist)){
        r <- rast(filelist[[i]])
        crs_2856 <- rast(target_rast)
        nameClean <- str_extract(filelist[[i]], pattern = "(?<=Export/)(.+)(?=.tif)")
    
        if(crs(r) != crs(crs_2856)){
            project(x = r, y = crs_2856, 
                    filename = paste0(filepath, "/",
                                              nameClean, "_2856", ".tif"),
                    overwrite = TRUE) 
        } else {
            next
        }
    }
}

tif_proc(filepath = filepath, target_rast = mas_dtw)

```


```{r}
mas_files <- list.files(filepath, pattern = ".tif$", full.names = TRUE) |> 
    # str_subset(pattern = "^(?!.*flow)") |> 
    # str_subset(pattern = "^(?!.*filled)") |> 
    # str_subset(pattern = "^(?!.*huc.tif)") |>
    str_subset(pattern = "2856.tif") 

mas_files
```


```{r}
mas_rasts <- c(rast(mas_files[c(-8,-9)]), rast(mas_dtw), rast("Mashel/Mashel_WIP_2024/DataExport/Mashel_DEM_HUC12_2856_4m_TWI.tif"), rast("Mashel/Mashel_WIP_2024/DataExport/Mashel_DEM_HUC12_2856_4m.tif"))

names(mas_rasts)
names(mas_rasts[[7]]) <- "CHM"
names(mas_rasts[[16]]) <- "DTW"
names(mas_rasts[[17]]) <- "TWI"
names(mas_rasts[[18]]) <- "DEM"
```




Training Points 
```{r}
mas_pts <- vect("Mashel/Mashel_WIP_2024/DataExport/Mashel_NWI_filter_train_pts.shp") |>
    project("EPSG:2856")
mas_pts
```

```{r}
mas_pts_train <- terra::extract(mas_rasts, mas_pts, bind = TRUE, filename = "Mashel/Mashel_WIP_2024/DataExport/MashelNWI_pts_training.gpkg")

mas_pts_train_df <- as.data.frame(mas_pts_train)
```




#### Random Forest Modeling
```{r eval=TRUE}
set.seed(11)

rf_mod_func <- function(points){
        train.index <- as.vector(sample(c(1:nrow(points)),
                                        0.7*nrow(points), replace=F))
        train <- points[train.index, ]
        test <- points[-train.index, ]
        
        rf_model <- randomForest(as.factor(class) ~ ., mtry = 10, 
                         sampsize = nrow(train[train$class == "WET",]),
                         replace = TRUE, #weights = wetwt, 
                         nodesize =1,
                         ntree = 1000, na.action = na.omit,
                         importance = TRUE, data = train)
        
        test_predict <- predict(rf_model, newdata = test, type = "response") 
        cm <- caret::confusionMatrix(test_predict, as.factor(test$class))
        return(list(test, rf_model, cm))
}
```


```{r eval=FALSE}
mas_model <- rf_mod_func(mas_pts_train_df)

mas_rf <- mas_model[[2]]
mas_test <- mas_model[[1]]
mas_cm <- mas_model[[3]]
mas_rf
mas_cm

```

Model output


#### WIP raster prediction
```{r}
#| eval: false
mas_pred_wip <- terra::predict(mas_rasts, mas_rf, type = "prob", 
                    filename = "Mashel/Mashel_WIP_2024/DataExport/Mashel_NWI_WIP_2856.tif", 
                    overwrite = TRUE)
```


```{r}
pred_wip <- rast("Mashel/Mashel_WIP_2024/DataExport/Mashel_NWI_WIP_2856.tif")

plot(pred_wip)
```

#### Stratified sample points across WIP

Can be adapted to do evenly spread across all of WIP or weighted to 1:2 WET:UPL for final validation
```{r}
#| eval: false


filepath <- "Mashel/Mashel_WIP_2024/DataExport"

strat_pts <- function(wip_rast){
    m <- c(0, 0.25, 1,
           0.25, 0.5, 2,
           0.5, 0.75, 3,
           0.75, 1, 3)
    rclmat <- matrix(m, ncol=3, byrow=TRUE)
    wip_rcl <- classify(wip_rast$WET, rclmat, include.lowest=TRUE)
    
    wip_stratified_points <- spatSample(wip_rcl, size = 50, 
                                        method = "stratified", 
                                        as.points = TRUE, 
                                        na.rm = TRUE)
    writeVector(wip_stratified_points, 
                paste0(filepath, "/", 
                       deparse(substitute(wip_rast)), 
                       "_NWI_stratified_points_validate_100wet", ".shp"),
                overwrite = TRUE)
    return(wip_stratified_points)
}

mas_strat_pts_valid <- strat_pts(wip_rast = pred_wip)


```


#### Bring stratified points back after reclassification in GIS

```{r}
mas_strat_pts_class <- vect("Mashel/Mashel_WIP_2024/DataExport/mas_pred_wip_NWI_stratified_600points_class.shp")

plot(mas_strat_pts_class, "validate", main = "Mashel User Classified Points")
```


Extract raster metrics again 

```{r}
mas_pts_train_2 <- mas_strat_pts_class |> select(validate) |> terra::extract(x = mas_rasts, bind = TRUE)
writeVector(mas_pts_train_2, "Mashel/Mashel_WIP_2024/DataExport/mas_strat_pts_class_extmetrics.shp",
            overwrite=TRUE)

mas_pts_validate <- vect("Mashel/Mashel_WIP_2024/DataExport/mas_pred_wip_NWI_stratified_points_validate.shp") |> select(validate) |> terra::extract(x = mas_rasts, bind = TRUE)
writeVector(mas_pts_validate, "Mashel/Mashel_WIP_2024/DataExport/mas_pred_wip_NWI_stratified_points_validate_extract.shp",
            overwrite = TRUE)

```


Rerun random forest algorithm 
```{r}
mas_pts_train_2_df <- vect("Mashel/Mashel_WIP_2024/DataExport/mas_strat_pts_class_extmetrics.shp") |> as.data.frame() |> dplyr::rename("class" = "validate")
mas_pts_validate_df <- vect("Mashel/Mashel_WIP_2024/DataExport/mas_pred_wip_NWI_stratified_points_validate_extract.shp") |> as.data.frame() |> dplyr::rename("class" = "validate")
```


```{r}
#| eval: false
filepath2 <- "Mashel/Mashel_WIP_2024/Model_Export"

rf_mod_func2 <- function(points, val_points, path){
    set.seed(11)
        #train.index <- points
        train <- points
        test <- val_points
        
        rf_model <- randomForest(as.factor(class) ~ ., mtry = 10, 
                         sampsize = nrow(train[train$class == "WET",]),
                         replace = TRUE,  
                         nodesize =1,
                         ntree = 1000, na.action = na.omit,
                         importance = TRUE, data = train)
        
        test_predict <- predict(rf_model, newdata = test, type = "response") 
        cm <- caret::confusionMatrix(test_predict, as.factor(test$class))
        
        save(rf_model, file = paste0(path, "/", 
                       toupper(str_sub(deparse(substitute(val_points)),1,3)), 
                       "_RF_Model_Final", ".RData"))
        
        return(list(test_predict, rf_model, cm))
}



mas_model_2 <- rf_mod_func2(mas_pts_train_2_df, mas_pts_validate_df, filepath2)
```


```{r}
mas_rf_2 <- mas_model_2[[2]]
mas_test_2 <- mas_model_2[[1]]
mas_cm_2 <- mas_model_2[[3]]
mas_rf_2
mas_cm_2
varImpPlot(mas_rf_2)
```


```{r}
mas_pred_wip_final <- terra::predict(mas_rasts, mas_rf_2, type = "prob", 
                    filename = "Mashel/Mashel_WIP_2024/DataExport/Mashel_WIP_Final_2856.tif", 
                    overwrite = TRUE)
```

```{r}
#| cache: true


mas_wip_final <- rast("Mashel/Mashel_WIP_2024/DataExport/Mashel_WIP_Final_2856.tif")[[2]]

pal <- colorNumeric("YlGnBu", values(mas_wip_final),
  na.color = "transparent")

leaflet::leaflet() |> 
    addProviderTiles(providers$Esri.WorldImagery) |> 
    addRasterImage(mas_wip_final, group = "Mashel WIP", colors = pal,
                   maxBytes = 20630008) |> 
    addLayersControl(overlayGroups = c("Mashel WIP")) |>
    addLegend(pal = pal, values = values(mas_pred_wip_final))
```


## Colville WIP 

```{r}
col_files <- list.files("Colville/Colville_WIP_2024/DataExport", pattern = ".tif$", full.names = TRUE) |> 
    str_subset(pattern = "^(?!.*flow)") |> 
    str_subset(pattern = "^(?!.*filled)") |> 
    str_subset(pattern = "^(?!.*huc.tif)") |>
    str_subset(pattern = "2855.tif") 

# for(i in 1:length(col_files)){
#     r <- rast(col_files[[i]])
#     crs_2855 <- crs(rast(col_files[[1]]))
#     nameClean <- str_extract(col_files[[i]], pattern = "(?<=Export/)(.+)(?=.tif)")
#     
#     if(crs(r) != crs_2855){
#         project(r, crs_2855, 
#                 filename = paste0("Colville/Colville_WIP_2024/DataExport/",
#                                           nameClean, "_2855", ".tif"))
#     } else {
#         next
#     }
# }

col_files
```


```{r}
col_twi <- rast("Colville/Colville_WIP_2024/DataExport/colvilledem_TWI.tif") 
col_dtw <- rast("Colville/Colville_WIP_2024/DataExport/Colville_DTW.tif")
names(col_twi) <- "TWI"
names(col_dtw) <- "DTW"
col_rasts <- c(rast(col_files), col_twi, col_dtw)
names(col_rasts[[3]]) <- "DEM"
```


```{r}
col_pts <- vect("Colville/Colville_WIP_2024/DataExport/ColvilleNWI_pts_merge.shp")
col_pts
```

```{r}
col_pts_train <- terra::extract(col_rasts, col_pts, bind = TRUE, filename = "Colville/Colville_WIP_2024/DataExport/ColvilleNWI_pts_merge_training.gpkg")

col_pts_train_df <- as.data.frame(col_pts_train)
```


```{r}
col_model <- rf_mod_func(col_pts_train_df)

col_rf <- col_model[[2]]
col_test <- col_model[[1]]
col_cm <- col_model[[3]]
col_rf
col_cm
```

#### NWI-based WIP model prediction to raster map

```{r}
#| eval: false

col_pred_wip <- predict(col_rasts, col_rf, type = "prob", 
                    filename = "Colville/Colville_WIP_2024/DataExport/Colville_NWI_WIP.tif", 
                    overwrite = TRUE)
```


```{r}
col_pred_wip <- rast("Colville/Colville_WIP_2024/DataExport/Colville_NWI_WIP.tif")
plot(col_pred_wip)
```

#### Stratify points across NWI-based WIP model prediction 



```{r}
filepath <- "Colville/Colville_WIP_2024/DataExport/"

strat_pts <- function(wip_rast){
    m <- c(0, 0.25, 1,
           0.25, 0.5, 2,
           0.5, 0.75, 3,
           0.75, 1, 4)
    rclmat <- matrix(m, ncol=3, byrow=TRUE)
    wip_rcl <- classify(wip_rast$WET, rclmat, include.lowest=TRUE)
    
    wip_stratified_points <- spatSample(wip_rcl, size = 150, 
                                        method = "stratified", 
                                        as.points = TRUE, 
                                        na.rm = TRUE,
                                        exp = 10)
    writeVector(wip_stratified_points, 
                paste0(filepath, 
                       deparse(substitute(wip_rast)), 
                       "_NWI_stratified_points_validate", ".shp"),
                overwrite = TRUE)
    return(wip_stratified_points)
}

col_strat_pts_valid <- strat_pts(wip_rast = col_pred_wip)

plot(col_strat_pts_valid, "WET")
```

Also do this to make a separate test dataset of 300 points (100 Wet 200 Upl)

```{r}
strat_pts_test <- function(wip_rast){
    set.seed(11)
    m <- c(0, 0.25, 1,
           0.25, 0.5, 2,
           0.5, 0.75, 3,
           0.75, 1, 3)
    rclmat <- matrix(m, ncol=3, byrow=TRUE)
    wip_rcl <- classify(wip_rast$WET, rclmat, include.lowest=TRUE)
    
    wip_stratified_points <- spatSample(wip_rcl, size = 100, 
                                        method = "stratified", 
                                        as.points = TRUE, 
                                        na.rm = TRUE,
                                        exp = 10)
    writeVector(wip_stratified_points, 
                paste0(filepath, 
                       deparse(substitute(wip_rast)), 
                       "_NWI_stratified_points_test", ".shp"),
                overwrite = TRUE)
    return(wip_stratified_points)
}

col_strat_pts_test <- strat_pts_test(wip_rast = col_pred_wip)
col_strat_pts_test
plot(col_strat_pts_test, "WET")
```


#### Bring stratified points back after reclassification in GIS

```{r}
col_strat_pts_test <- vect("Colville/Colville_WIP_2024/DataExport/col_pred_wip_NWI_stratified_points_validate.shp")

plot(col_strat_pts_test, "validate", main = "Colville User Classified Points")
```


Extract raster metrics again 

```{r}
col_pts_train_2 <- col_strat_pts_class |> dplyr::select(validate) |> terra::extract(x = col_rasts, bind = TRUE)
writeVector(col_pts_train_2, "Colville/Colville_WIP_2024/DataExport/col_strat_pts_class_extmetrics.shp",
            overwrite=TRUE)

col_pts_test <- vect("Colville/Colville_WIP_2024/DataExport/col_pred_wip_NWI_stratified_points_validate.shp") |> dplyr::select(validate) |> terra::extract(x = col_rasts, bind = TRUE)
writeVector(col_pts_test, "Colville/Colville_WIP_2024/DataExport/col_pred_wip_NWI_stratified_points_validate_extract.shp",
            overwrite = TRUE)

```


Rerun random forest algorithm 
```{r}
mas_pts_train_2_df <- vect("Mashel/Mashel_WIP_2024/DataExport/mas_strat_pts_class_extmetrics.shp") |> as.data.frame() |> dplyr::rename("class" = "validate")
mas_pts_validate_df <- vect("Mashel/Mashel_WIP_2024/DataExport/mas_pred_wip_NWI_stratified_points_validate_extract.shp") |> as.data.frame() |> dplyr::rename("class" = "validate")
```


```{r}
#| eval: false
filepath2 <- "Mashel/Mashel_WIP_2024/Model_Export"

rf_mod_func2 <- function(points, val_points, path){
    set.seed(11)
        #train.index <- points
        train <- points
        test <- val_points
        
        rf_model <- randomForest(as.factor(class) ~ ., mtry = 10, 
                         sampsize = nrow(train[train$class == "WET",]),
                         replace = TRUE,  
                         nodesize =1,
                         ntree = 1000, na.action = na.omit,
                         importance = TRUE, data = train)
        
        test_predict <- predict(rf_model, newdata = test, type = "response") 
        cm <- caret::confusionMatrix(test_predict, as.factor(test$class))
        
        save(rf_model, file = paste0(path, "/", 
                       toupper(str_sub(deparse(substitute(val_points)),1,3)), 
                       "_RF_Model_Final", ".RData"))
        
        return(list(test_predict, rf_model, cm))
}



mas_model_2 <- rf_mod_func2(mas_pts_train_2_df, mas_pts_validate_df, filepath2)
```


```{r}
mas_rf_2 <- mas_model_2[[2]]
mas_test_2 <- mas_model_2[[1]]
mas_cm_2 <- mas_model_2[[3]]
mas_rf_2
mas_cm_2
varImpPlot(mas_rf_2)
```